---
title: "PKTu"
subtitle: "Gene Set Enrichment Analysis (GSEA)"
author: "Lachlan Baer"
date-modified: today
date-format: "DD MMMM, YYYY"
modified-title: "LAST MODIFIED"
code-fold: true
fig-width: 8
fig-height: 6
fig-align: center
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    error = FALSE,
    fig.align = "center"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r pkgs}
suppressPackageStartupMessages({
    ## Common packages
    library(tidyverse)
    library(magrittr)
    library(here)
    library(kableExtra)
    library(RColorBrewer)
    library(ggpubr)
    library(scales)
    library(AnnotationHub)
    library(ggrepel)
    library(ggtext)
    library(glue)
    library(pander)
    library(DT)
    library(reactable)
    library(reactablefmtr)
    library(htmltools)
    ## Document-specific packages
    library(msigdbr)
    library(fgsea)
    library(limma)
    library(goseq)
    library(ggraph)
    library(tidygraph)
    library(ComplexHeatmap)
})
```

```{r opts}
theme_set(theme_bw())
theme_update(
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown(),
    strip.background = element_blank(),
    strip.text = element_textbox(
        size = 10,
        color = "black", fill = "white", box.color = "black",
        halign = 0.5, linetype = 1, width = unit(1, "npc"),
        padding = margin(3, 0, 1, 0), margin = margin(0, 0, 0, 0)
    )
)
pal1 <- brewer.pal(12, "Set3")
pal2 <- brewer.pal(8, "Set2")
```

```{r pal}
pal_frax_fmr1 <- c(
    "Not DE" = "black", "FRAX only" = pal[3], "FMR1 only" = pal[2], "FRAX & FMR1" = pal[4]
)
```

```{r lb_complex_dt}
lb_complex_dt <- function(tbl, cap, rownames = FALSE){
    datatable(
        tbl,
        caption = htmltools::tags$caption(
            cap,
            style = "text-align: left;"
        ),
        options = list(
            dom = "tflipr",  # Available options: tflipr
            columnDefs = list(list(className = 'dt-left', targets = "_all")),  # Align all columns left
            pageLength = 10,  # Show all results
            scrollCollapse = TRUE,  # Removes whitespace when table is small
            scrollY = "100%",  # Sets max height of table before scrollbar is added
            scrollX = TRUE,  # Enable horizontal scrolling when too wide
            sScrollX = "100%",  # Stops last column title being cut off
            initComplete = JS(  # Set font
                "function(settings, json) {",
                "$('body').css({'font-family': 'Calibri'});",
                "}"
            )
        ),
        rownames = rownames,
        height = "100%",  # Seems to work in tandem with scrollCollapse
        width = "100%",  # Fills width when viewed in RStudio viewer
        class = "stripe hover row-border order-column compact nowrap"
    )
}
```

```{r lb_reactable}
lb_reactable <- function(
        tbl, highlight = TRUE, striped = TRUE, compact = TRUE,
        wrap = FALSE, resizable = TRUE, searchable = TRUE,
        style = list(fontFamily = "Calibri, sans-serif"), ...
){
    reactable(
        tbl,
        highlight = highlight, striped = striped, compact = compact,
        wrap = wrap, resizable = TRUE, searchable = TRUE,
        style = style, ...
    )
}
tooltip <- function(value, tooltip) {
    tags$abbr(
        style = "text-decoration: underline; text-decoration-style: solid; cursor: help",
        title = tooltip,
        value
    )
}
react_format <- function(format, digits = 2){
    function(val, ind, col_name){
        formatC(val, digits = digits, format = format)
    }
}
```

# Metadata

```{r meta}
meta <- read_csv(here("misc/metadata.csv")) %>%
    mutate(
        geno_psen1 = fct_relevel(geno_psen1, c("wt", "eofad", "fai")),
        geno_chr14 = fct_relevel(geno_chr14, c("tu", "pk", "het")),
        geno_joint = fct_relevel(geno_joint, unique(geno_joint))
    )
meta %>%
    kable(
        align = "l",
        caption = "Sample metadata"
    ) %>%
    kable_styling(
        bootstrap_options = c("striped", "hover", "condensed", "responsive")
    )
```

```{r pal_group}
pal_group <- c(
    "eofad_tu" = pal1[1], "eofad_pk" = pal1[2], "eofad_het" = pal1[3],
    "fai_tu" = pal1[4], "fai_pk" = pal1[5], "fai_het" = pal1[6],
    "wt_tu" = pal1[7], "wt_pk" = pal1[8], "wt_het" = pal1[10],
    "eofad" = pal2[1], "fai" = pal2[2], "wt" = pal2[3],
    "pk" = pal2[4], "tu" = pal2[5], "het" = pal2[6]
) 
```

# DE data

Load data from the preceeding DE analysis

```{r load_data}
dge_list <- readRDS(here("Rdata/dge_list.Rds"))
tt_psen1 <- readRDS(here("Rdata/tt_psen1.Rds"))
tt_chr14 <- readRDS(here("Rdata/tt_chr14.Rds"))
tt_psen1_cqn <- readRDS(here("Rdata/tt_psen1_cqn.Rds"))
tt_chr14_cqn <- readRDS(here("Rdata/tt_chr14_cqn.Rds"))
```

# Annotations

```{r ensParams}
ens_species <- "Homo sapiens"
ens_release <- "112"
ens_assembly <- "GRCh38"
```

```{r ensDb}
ah <- AnnotationHub() %>%
    subset(species == ens_species) %>%
    subset(rdataclass == "EnsDb")
ahId <- ah$ah_id[str_detect(ah$title, ens_release)]
ensDb <- ah[[ahId]]
```

```{r primary_chrs}
primary_chrs <- c(as.character(1:22), "X", "Y", "MT")
```

```{r transcripts}
transcripts <- transcripts(ensDb)
tx_len <- exonsBy(ensDb, "tx") %>%
    width() %>%
    lapply(sum) %>%
    unlist(use.names = TRUE) %>%
    enframe(name = "tx_id", value = "length")
mcols(transcripts) <- mcols(transcripts)[
    c("tx_id", "gene_id", "gc_content")
] %>%
    as.data.frame() %>%
    left_join(tx_len)
```

```{r gene_gclen}
gene_gclen <- transcripts %>%
    mcols() %>%
    as_tibble() %>%
    group_by(gene_id) %>%
    summarise(
        gc_content = sum(gc_content*length) / sum(length),
        length = ceiling(median(length))
    )
```

```{r genes}
genes <- genes(ensDb)
mcols(genes) <- mcols(genes)[
    c("gene_id", "gene_name", "gene_biotype", "entrezid")
] %>%
    as.data.frame() %>%
    left_join(gene_gclen)
```

An `EnsDb` object was obtained for Ensembl release `r ens_release` with the `AnnotationHub` package.
This contained the information required to set up gene and transcript annotations.
Gene-level estimates of GC content and length were also generated which are required for assessing potential biases in downstream analysis.

# Gene sets

Multiple gene set databases were accessed with the `msigdbr` package.
For this analysis the Hallmark, KEGG and Wikipathways gene sets were chosen.

The `babelgene` and `msigdbr` package versions dictate the resulting gene sets, and therefore the gene sets were saved in the project directory after initial loading.
This avoids reproducibility problems when re-running the analysis in the future.
The gene sets in this analysis were obtained using `babelgene v22.9` and `msigdbr v7.5.1`.

```{r hm}
hm_path <- here("Rdata/hm.csv")
if (!file.exists(hm_path)) {
    hm <- msigdbr(ens_species, category = "H") %>%
        dplyr::filter(!is.na(ensembl_gene)) %>%
        distinct(gs_name, ensembl_gene, .keep_all = TRUE) %>%
        mutate(gs_name = str_remove(gs_name, "^HALLMARK_"))
    write_csv(hm, hm_path)
} else {
    hm <- read_csv(hm_path)
}
hm_by_gene <- hm %>%
    split(f = .$ensembl_gene) %>%
    lapply(extract2, "gs_name")
hm_by_id <- hm %>%
    split(f = .$gs_name) %>%
    lapply(extract2, "ensembl_gene")
```

The Hallmark gene sets consists of `r length(hm_by_id)` gene sets ranging from sizes of `r min(sapply(hm_by_id, length))` to `r max(sapply(hm_by_id, length))` genes.

```{r kg}
kg_path <- here("Rdata/kg.csv")
if (!file.exists(kg_path)) {
    kg <- msigdbr(ens_species, category = "C2", subcategory = "CP:KEGG") %>%
        dplyr::filter(!is.na(ensembl_gene)) %>%
        distinct(gs_name, ensembl_gene, .keep_all = TRUE) %>%
        mutate(gs_name = str_remove(gs_name, "^KEGG_"))
    write_csv(kg, kg_path)
} else {
    kg <- read_csv(kg_path)
}
kg_by_gene <- kg %>%
    split(f = .$ensembl_gene) %>%
    lapply(extract2, "gs_name")
kg_by_id <- kg %>%
    split(f = .$gs_name) %>%
    lapply(extract2, "ensembl_gene")
```

The KEGG gene sets consists of `r length(kg_by_id)` gene sets ranging from sizes of `r min(sapply(kg_by_id, length))` to `r max(sapply(kg_by_id, length))` genes.

```{r wk}
wk_path <- here("Rdata/wk.csv")
if (!file.exists(wk_path)) {
    wk <- msigdbr(ens_species, category = "C2", subcategory = "CP:WIKIPATHWAYS") %>%
        dplyr::filter(!is.na(ensembl_gene)) %>%
        distinct(gs_name, ensembl_gene, .keep_all = TRUE) %>%
        mutate(gs_name = str_remove(gs_name, "^WP_"))
    write_csv(wk, wk_path)
} else {
    wk <- read_csv(wk_path)
}
wk_by_gene <- wk %>%
    split(f = .$ensembl_gene) %>%
    lapply(extract2, "gs_name")
wk_by_id <- wk %>%
    split(f = .$gs_name) %>%
    lapply(extract2, "ensembl_gene")
```

The Wikipathways gene sets consists of `r length(wk_by_id)` gene sets ranging from sizes of `r min(sapply(wk_by_id, length))` to `r max(sapply(wk_by_id, length))` genes.

```{r rtm}
rtm_path <- here("Rdata/rtm.csv")
if (!file.exists(rtm_path)) {
    rtm <- msigdbr(ens_species, category = "C2", subcategory = "CP:REACTOME") %>%
        dplyr::filter(!is.na(ensembl_gene)) %>%
        distinct(gs_name, ensembl_gene, .keep_all = TRUE) %>%
        mutate(gs_name = str_remove(gs_name, "^REACTOME_"))
    write_csv(rtm, rtm_path)
} else {
    rtm <- read_csv(rtm_path)
}
rtm_by_gene <- rtm %>%
    split(f = .$ensembl_gene) %>%
    lapply(extract2, "gs_name")
rtm_by_id <- rtm %>%
    split(f = .$gs_name) %>%
    lapply(extract2, "ensembl_gene")
```

The Reactome gene sets consists of `r length(rtm_by_id)` gene sets ranging from sizes of `r min(sapply(rtm_by_id, length))` to `r max(sapply(rtm_by_id, length))` genes.

# GSEA

```{r gsea_node_network}
gsea_node_network <- function(gsea, directional = TRUE, debug = FALSE){
    
    if (debug) browser()
    
    sig <- gsea %>%
        dplyr::filter(Sig)
    
    colours <- hue_pal()(nrow(sig)) %>%
        set_names(sig$Pathway) %>%
        c(
            "None" = "black",
            "Up" = "darkblue",
            "Down" = "darkred"
        )
    
    p_net <- sig %>%
        dplyr::select(Pathway, leadingEdge) %>%
        mutate(
            leadingEdge = lapply(leadingEdge, \(x){
                sort(genes$gene_name[genes$gene_id %in% x])
            })
        ) %>%
        unnest(cols = "leadingEdge") %>%
        as_tbl_graph() %>%
        activate("nodes") %>%
        mutate(
            Edge = vapply(
                name,
                \(x){
                    ifelse(
                        x %in% gsea$Pathway,
                        gsea$Edge[gsea$Pathway == x],
                        1
                    )
                },
                numeric(1)
            ),
            pval = vapply(
                name,
                \(x){
                    ifelse(
                        x %in% gsea$Pathway,
                        gsea$pval[gsea$Pathway == x],
                        1
                    )
                },
                numeric(1)
            ),
            is_pathway = name %in% gsea$Pathway,
            pathway = ifelse(is_pathway, name, "None")
        )
    
    if (directional) {
        p_net <- p_net %>%
            mutate(
                Direction = vapply(
                    name,
                    \(x){
                        ifelse(
                            x %in% gsea$Pathway,
                            gsea$Direction[gsea$Pathway == x],
                            ""
                        )
                    },
                    character(1)
                )
            )
    } else {
        p_net <- p_net %>%
            mutate(Direction = "None")
    }
    
    p_net <- p_net %>%
        activate("edges") %>%
        mutate(
            from_pathway = gsea$Pathway[from]
        ) %>%
        ggraph(layout = "kk") +
        geom_edge_link(aes(colour = from_pathway), alpha = 0.2) +
        geom_node_point(aes(size = -log10(pval), fill = pathway), colour = "black", shape = 21) +
        geom_node_text(
            aes(label = name, colour = Direction),
            data = . %>% 
                dplyr::filter(name %in% gsea$Pathway) %>%
                mutate(
                    name = str_replace_all(name, "_", " ") %>%
                        str_trunc(width = 40) %>%
                        str_wrap(width = 20)
                ),
            size = 3, repel = TRUE, bg.color = "white", bg.r = 0.1
        ) +
        scale_colour_manual(values = colours) +
        scale_fill_manual(values = colours) +
        scale_edge_colour_manual(values = colours) +
        theme_void() +
        guides(edge_alpha = "none", edge_width = "none") +
        scale_x_continuous(expand = expansion(c(0.05, 0.05))) +
        scale_y_continuous(expand = expansion(c(0.05, 0.05))) +
        theme(legend.position = "none")
    
    p_net
    
}
```

```{r gsea_edge_network}
gsea_edge_network <- function(gsea, directional = TRUE, debug = FALSE){
    
    if (debug) browser()
    
    sig <- gsea %>%
        dplyr::filter(Sig)
    
    colours <- hue_pal()(nrow(sig)) %>%
        set_names(sig$Pathway) %>%
        c(
            "None" = "black",
            "Up" = "darkblue",
            "Down" = "darkred"
        )
    
    p_net <- lapply(seq_along(sig$Pathway), \(i){
        path <- sig$Pathway[i]
        le_path <- unlist(sig$leadingEdge[sig$Pathway == path])
        remaining <- sig$Pathway[(i):length(sig$Pathway)]
        lapply(remaining, \(x){
            le_x <- unlist(sig$leadingEdge[sig$Pathway == x])
            overlap <- length(intersect(le_path, le_x))
            tibble(from = path, to = x, n = overlap)
        }) %>%
            bind_rows()
    }) %>%
        bind_rows() %>%
        dplyr::filter(n > 0) %>%
        as_tbl_graph() %>%
        activate("nodes") %>%
        mutate(
            pval = vapply(
                name,
                \(x){
                    ifelse(
                        x %in% gsea$Pathway,
                        gsea$pval[gsea$Pathway == x],
                        1
                    )
                },
                numeric(1)
            )
        )
    
    if (directional) {
        p_net <- p_net %>%
            mutate(
                Direction = vapply(
                    name, \(x){gsea$Direction[gsea$Pathway == x]}, character(1)
                )
            )
    } else {
        p_net <- p_net %>%
            mutate(Direction = "None")
    }
    
    p_net <- p_net %>%
        ggraph(layout = "kk") +
        geom_edge_link(aes(alpha = n)) +
        geom_node_point(aes(size = -log10(pval), fill = name), colour = "black", shape = 21) +
        geom_node_text(
            aes(label = name, colour = Direction),
            data = . %>% 
                mutate(
                    name = str_replace_all(name, "_", " ") %>%
                        str_trunc(width = 30)
                ),
            size = 2.5, repel = TRUE, bg.color = "white", bg.r = 0.1
        ) +
        scale_colour_manual(values = colours) +
        scale_fill_manual(values = colours) +
        theme_void() +
        guides(edge_alpha = "none", edge_width = "none") +
        scale_x_continuous(expand = expansion(c(0.05, 0.05))) +
        scale_y_continuous(expand = expansion(c(0.05, 0.05))) +
        theme(legend.position = "none")
    
    p_net
    
}
```

## Directional GSEA

Directional GSEA was performed using a ranked list of genes ordered by significance of DE analysis.
Here the most significantly upregulated genes are found at one end of the list, while the most significantly down-regulated genes are found at the other.

### psen1

```{r ranks_d_psen1}
ranks_d_psen1 <- lapply(tt_psen1, \(x){
    x %>%
        mutate(stat = -log10(PValue) * sign(logFC)) %>%
        dplyr::arrange(desc(stat)) %>%
        with(structure(stat, names = gene_id))
})
```

#### Hallmark

```{r gsea_hm_d}
set.seed(2407)
gsea_hm_d <- lapply(ranks_d, \(x){
    fgseaMultilevel(hm_by_id, x) %>%
        as_tibble() %>%
        dplyr::rename(FDR = padj, Pathway = pathway) %>%
        dplyr::arrange(pval) %>%
        mutate(
            Edge = vapply(leadingEdge, length, numeric(1)),
            Direction = case_when(
                sign(ES) == 1 ~ "Up",
                sign(ES) == -1 ~ "Down",
                sign(ES) == 0 ~ "None"
            ),
            bonf.padj = p.adjust(pval, "bonferroni"),
            Sig = FDR < 0.05,
        )
})
```

##### FRAX

::: {.panel-tabset group="directional"}

###### Results table

```{r gsea_hm_d_frax}
id <- "gsea_hm_d_frax"
tbl <- gsea_hm_d$FRAX %>%
    dplyr::select(Pathway, pval, FDR, Edge, Direction, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

###### Network vis

```{r node_hm_d_frax}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_hm_d$FRAX, TRUE)
```

```{r edge_hm_d_frax}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_hm_d$FRAX, TRUE)
```

:::

##### FMR1

::: {.panel-tabset group="directional"}

###### Results table

```{r gsea_hm_d_fmr1}
id <- "gsea_hm_d_fmr1"
tbl <- gsea_hm_d$FMR1 %>%
    dplyr::select(Pathway, pval, FDR, Edge, Direction, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

###### Network vis

```{r node_hm_d_fmr1}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_hm_d$FMR1, TRUE)
```

```{r edge_hm_d_fmr1}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_hm_d$FMR1, TRUE)
```

:::

##### Comparison

```{r heat_hm_d}
tbl <- gsea_hm_d$FRAX %>%
    dplyr::select(Pathway, FRAX = NES, sig_frax = Sig) %>%
    left_join(gsea_hm_d$FMR1) %>%
    dplyr::select(Pathway, FRAX, FMR1 = NES, sig_frax, sig_fmr1 = Sig) %>%
    mutate(
        Pathway = str_trunc(str_replace_all(Pathway, "_", " "), 30),
        sig_any = as.logical(pmax(sig_frax, sig_fmr1))
    ) %>%
    dplyr::filter(sig_any)
mat <- tbl %>%
    dplyr::select(Pathway, FRAX, FMR1) %>%
    column_to_rownames("Pathway") %>%
    as.matrix()
row_anno <-  rowAnnotation(
    Significant = tbl %>%
        mutate(
            category = case_when(
                sig_frax & sig_fmr1 ~ "FRAX & FMR1",
                sig_frax & !sig_fmr1 ~ "FRAX only",
                !sig_frax & sig_fmr1 ~ "FMR1 only"
            ),
            category = fct_relevel(category, c(
                "FRAX only",
                "FMR1 only",
                "FRAX & FMR1"
            ))
        ) %>%
        with(structure(category, names = Pathway)),
    col = list(
        Significant = pal_frax_fmr1
    )
)
Heatmap(
    mat,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 10),
    row_names_side = "left",
    left_annotation = row_anno,
    col = colorRampPalette(c("darkred", "lightyellow", "darkblue"))(100),
    heatmap_legend_param = list(
        title = "NES"
    )
)
```

### KEGG

```{r gsea_kg_d}
set.seed(2407)
gsea_kg_d <- lapply(ranks_d, \(x){
    fgseaMultilevel(kg_by_id, x) %>%
        as_tibble() %>%
        dplyr::rename(FDR = padj, Pathway = pathway) %>%
        dplyr::arrange(pval) %>%
        mutate(
            Edge = vapply(leadingEdge, length, numeric(1)),
            Direction = case_when(
                sign(ES) == 1 ~ "Up",
                sign(ES) == -1 ~ "Down",
                sign(ES) == 0 ~ "None"
            ),
            bonf.padj = p.adjust(pval, "bonferroni"),
            Sig = FDR < 0.05,
        )
})
```

#### FRAX

::: {.panel-tabset group="directional"}

##### Results table

```{r gsea_kg_d_frax}
id <- "gsea_kg_d_frax"
tbl <- gsea_kg_d$FRAX %>%
    dplyr::select(Pathway, pval, FDR, Edge, Direction, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_kg_d_frax}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_kg_d$FRAX, TRUE)
```

```{r edge_kg_d_frax}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_kg_d$FRAX, TRUE)
```

:::

#### FMR1

::: {.panel-tabset group="directional"}

##### Results table

```{r gsea_kg_d_fmr1}
id <- "gsea_kg_d_fmr1"
tbl <- gsea_kg_d$FMR1 %>%
    dplyr::select(Pathway, pval, FDR, Edge, Direction, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_kg_d_fmr1}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_kg_d$FMR1, TRUE)
```

```{r edge_kg_d_fmr1}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_kg_d$FMR1, TRUE)
```

:::

#### Comparison

```{r heat_kg_d}
tbl <- gsea_kg_d$FRAX %>%
    dplyr::select(Pathway, FRAX = NES, sig_frax = Sig) %>%
    left_join(gsea_kg_d$FMR1) %>%
    dplyr::select(Pathway, FRAX, FMR1 = NES, sig_frax, sig_fmr1 = Sig) %>%
    mutate(
        Pathway = str_trunc(str_replace_all(Pathway, "_", " "), 30),
        sig_any = as.logical(pmax(sig_frax, sig_fmr1))
    ) %>%
    dplyr::filter(sig_any)
mat <- tbl %>%
    dplyr::select(Pathway, FRAX, FMR1) %>%
    column_to_rownames("Pathway") %>%
    as.matrix()
row_anno <-  rowAnnotation(
    Significant = tbl %>%
        mutate(
            category = case_when(
                sig_frax & sig_fmr1 ~ "FRAX & FMR1",
                sig_frax & !sig_fmr1 ~ "FRAX only",
                !sig_frax & sig_fmr1 ~ "FMR1 only"
            ),
            category = fct_relevel(category, c(
                "FRAX only",
                "FMR1 only",
                "FRAX & FMR1"
            ))
        ) %>%
        with(structure(category, names = Pathway)),
    col = list(
        Significant = pal_frax_fmr1
    )
)
Heatmap(
    mat,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 10),
    row_names_side = "left",
    left_annotation = row_anno,
    col = colorRampPalette(c("darkred", "lightyellow", "darkblue"))(100),
    heatmap_legend_param = list(
        title = "NES"
    )
)
```

### Wikipathways

```{r gsea_wk_d}
set.seed(2407)
gsea_wk_d <- lapply(ranks_d, \(x){
    fgseaMultilevel(wk_by_id, x) %>%
        as_tibble() %>%
        dplyr::rename(FDR = padj, Pathway = pathway) %>%
        dplyr::arrange(pval) %>%
        mutate(
            Edge = vapply(leadingEdge, length, numeric(1)),
            Direction = case_when(
                sign(ES) == 1 ~ "Up",
                sign(ES) == -1 ~ "Down",
                sign(ES) == 0 ~ "None"
            ),
            bonf.padj = p.adjust(pval, "bonferroni"),
            Sig = FDR < 0.05,
        )
})
```

#### FRAX

::: {.panel-tabset group="directional"}

##### Results table

```{r gsea_wk_d_frax}
id <- "gsea_wk_d_frax"
tbl <- gsea_wk_d$FRAX %>%
    dplyr::select(Pathway, pval, FDR, Edge, Direction, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_wk_d_frax}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_wk_d$FRAX, TRUE)
```

```{r edge_wk_d_frax}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_wk_d$FRAX, TRUE)
```

:::

#### FMR1

::: {.panel-tabset group="directional"}

##### Results table

```{r gsea_wk_d_fmr1}
id <- "gsea_wk_d_fmr1"
tbl <- gsea_wk_d$FMR1 %>%
    dplyr::select(Pathway, pval, FDR, Edge, Direction, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_wk_d_fmr1}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_wk_d$FMR1, TRUE)
```

```{r edge_wk_d_fmr1}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_wk_d$FMR1, TRUE)
```

:::

#### Comparison

```{r heat_wk_d}
tbl <- gsea_wk_d$FRAX %>%
    dplyr::select(Pathway, FRAX = NES, sig_frax = Sig) %>%
    left_join(gsea_wk_d$FMR1) %>%
    dplyr::select(Pathway, FRAX, FMR1 = NES, sig_frax, sig_fmr1 = Sig) %>%
    mutate(
        Pathway = str_trunc(str_replace_all(Pathway, "_", " "), 30),
        sig_any = as.logical(pmax(sig_frax, sig_fmr1))
    ) %>%
    dplyr::filter(sig_any)
mat <- tbl %>%
    dplyr::select(Pathway, FRAX, FMR1) %>%
    column_to_rownames("Pathway") %>%
    as.matrix()
row_anno <-  rowAnnotation(
    Significant = tbl %>%
        mutate(
            category = case_when(
                sig_frax & sig_fmr1 ~ "FRAX & FMR1",
                sig_frax & !sig_fmr1 ~ "FRAX only",
                !sig_frax & sig_fmr1 ~ "FMR1 only"
            ),
            category = fct_relevel(category, c(
                "FRAX only",
                "FMR1 only",
                "FRAX & FMR1"
            ))
        ) %>%
        with(structure(category, names = Pathway)),
    col = list(
        Significant = pal_frax_fmr1
    )
)
Heatmap(
    mat,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 8),
    row_names_side = "left",
    left_annotation = row_anno,
    col = colorRampPalette(c("darkred", "lightyellow", "darkblue"))(100),
    heatmap_legend_param = list(
        title = "NES"
    )
)
```

### Reactome

```{r gsea_rtm_d}
set.seed(2407)
gsea_rtm_d <- lapply(ranks_d, \(x){
    fgseaMultilevel(rtm_by_id, x) %>%
        as_tibble() %>%
        dplyr::rename(FDR = padj, Pathway = pathway) %>%
        dplyr::arrange(pval) %>%
        mutate(
            Edge = vapply(leadingEdge, length, numeric(1)),
            Direction = case_when(
                sign(ES) == 1 ~ "Up",
                sign(ES) == -1 ~ "Down",
                sign(ES) == 0 ~ "None"
            ),
            bonf.padj = p.adjust(pval, "bonferroni"),
            Sig = FDR < 0.05,
        )
})
```

#### FRAX

::: {.panel-tabset group="directional"}

##### Results table

```{r gsea_rtm_d_frax}
id <- "gsea_rtm_d_frax"
tbl <- gsea_rtm_d$FRAX %>%
    dplyr::select(Pathway, pval, FDR, Edge, Direction, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_rtm_d_frax}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_rtm_d$FRAX, TRUE)
```

```{r edge_rtm_d_frax}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_rtm_d$FRAX, TRUE)
```

:::

#### FMR1

::: {.panel-tabset group="directional"}

##### Results table

```{r gsea_rtm_d_fmr1}
id <- "gsea_rtm_d_fmr1"
tbl <- gsea_rtm_d$FMR1 %>%
    dplyr::select(Pathway, pval, FDR, Edge, Direction, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_rtm_d_fmr1}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_rtm_d$FMR1, TRUE)
```

```{r edge_rtm_d_fmr1}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_rtm_d$FMR1, TRUE)
```

:::

#### Comparison

```{r heat_rtm_d}
#| fig-height: 12
tbl <- gsea_rtm_d$FRAX %>%
    dplyr::select(Pathway, FRAX = NES, sig_frax = Sig) %>%
    left_join(gsea_rtm_d$FMR1) %>%
    dplyr::select(Pathway, FRAX, FMR1 = NES, sig_frax, sig_fmr1 = Sig) %>%
    mutate(
        Pathway = str_trunc(str_replace_all(Pathway, "_", " "), 30),
        sig_any = as.logical(pmax(sig_frax, sig_fmr1))
    ) %>%
    dplyr::filter(sig_any)
mat <- tbl %>%
    dplyr::select(Pathway, FRAX, FMR1) %>%
    column_to_rownames("Pathway") %>%
    as.matrix()
row_anno <-  rowAnnotation(
    Significant = tbl %>%
        mutate(
            category = case_when(
                sig_frax & sig_fmr1 ~ "FRAX & FMR1",
                sig_frax & !sig_fmr1 ~ "FRAX only",
                !sig_frax & sig_fmr1 ~ "FMR1 only"
            ),
            category = fct_relevel(category, c(
                "FRAX only",
                "FMR1 only",
                "FRAX & FMR1"
            ))
        ) %>%
        with(structure(category, names = Pathway)),
    col = list(
        Significant = pal_frax_fmr1
    )
)
Heatmap(
    mat,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 5),
    row_names_side = "left",
    left_annotation = row_anno,
    col = colorRampPalette(c("darkred", "lightyellow", "darkblue"))(100),
    heatmap_legend_param = list(
        title = "NES"
    )
)
```

## Non-directional GSEA

GSEA was also performed non-directionally, where the direction of expression change was not taken into account.
This can be useful because changes in transcript concentrations due to a certain functional response may correlate poorly with protein concentrations.
Even when the correlation is high, the activity of a protein may be modulated by post-translational modifications.
Non-directional GSEA provides an alternative perspective to our data.

```{r ranks_nd}
ranks_nd <- lapply(tt, \(x){
    x %>%
        mutate(stat = -log10(PValue)) %>%
        dplyr::arrange(desc(stat)) %>%
        with(structure(stat, names = gene_id))
})
```

### Hallmark

```{r gsea_hm_nd}
set.seed(2407)
gsea_hm_nd <- lapply(ranks_nd, \(x){
    fgseaMultilevel(hm_by_id, x, scoreType = "pos") %>%
        as_tibble() %>%
        dplyr::rename(FDR = padj, Pathway = pathway) %>%
        dplyr::arrange(pval) %>%
        mutate(
            Edge = vapply(leadingEdge, length, numeric(1)),
            bonf.padj = p.adjust(pval, "bonferroni"),
            Sig = FDR < 0.05,
        )
})
```

#### FRAX

::: {.panel-tabset group="nondirectional"}

##### Results tables

```{r gsea_hm_nd_frax}
id <- "gsea_hm_nd_frax"
tbl <- gsea_hm_nd$FRAX %>%
    dplyr::select(Pathway, pval, FDR, Edge, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_hm_nd_frax}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_hm_nd$FRAX, FALSE)
```

```{r edge_hm_nd_frax}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_hm_nd$FRAX, FALSE)
```

:::

#### FMR1

::: {.panel-tabset group="nondirectional"}

##### Results tables

```{r gsea_hm_nd_fmr1}
id <- "gsea_hm_nd_fmr1"
tbl <- gsea_hm_nd$FMR1 %>%
    dplyr::select(Pathway, pval, FDR, Edge, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_hm_nd_fmr1}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_hm_nd$FMR1, FALSE)
```

```{r edge_hm_nd_fmr1}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_hm_nd$FMR1, FALSE)
```

:::

#### Comparison

```{r heat_hm_nd}
tbl <- gsea_hm_nd$FRAX %>%
    dplyr::select(Pathway, FRAX = NES, sig_frax = Sig) %>%
    left_join(gsea_hm_nd$FMR1) %>%
    dplyr::select(Pathway, FRAX, FMR1 = NES, sig_frax, sig_fmr1 = Sig) %>%
    mutate(
        Pathway = str_trunc(str_replace_all(Pathway, "_", " "), 30),
        sig_any = as.logical(pmax(sig_frax, sig_fmr1))
    ) %>%
    dplyr::filter(sig_any)
mat <- tbl %>%
    dplyr::select(Pathway, FRAX, FMR1) %>%
    column_to_rownames("Pathway") %>%
    as.matrix()
row_anno <-  rowAnnotation(
    Significant = tbl %>%
        mutate(
            category = case_when(
                sig_frax & sig_fmr1 ~ "FRAX & FMR1",
                sig_frax & !sig_fmr1 ~ "FRAX only",
                !sig_frax & sig_fmr1 ~ "FMR1 only"
            ),
            category = fct_relevel(category, c(
                "FRAX only",
                "FMR1 only",
                "FRAX & FMR1"
            ))
        ) %>%
        with(structure(category, names = Pathway)),
    col = list(
        Significant = pal_frax_fmr1
    )
)
Heatmap(
    mat,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 10),
    row_names_side = "left",
    left_annotation = row_anno,
    col = colorRampPalette(c("grey90", "darkblue"))(100),
    heatmap_legend_param = list(
        title = "NES"
    )
)
```

### KEGG

```{r gsea_kg_nd}
set.seed(2407)
gsea_kg_nd <- lapply(ranks_nd, \(x){
    fgseaMultilevel(kg_by_id, x, scoreType = "pos") %>%
        as_tibble() %>%
        dplyr::rename(FDR = padj, Pathway = pathway) %>%
        dplyr::arrange(pval) %>%
        mutate(
            Edge = vapply(leadingEdge, length, numeric(1)),
            bonf.padj = p.adjust(pval, "bonferroni"),
            Sig = FDR < 0.05,
        )
})
```

#### FRAX

::: {.panel-tabset group="nondirectional"}
##### Results tables

```{r gsea_kg_nd_frax}
id <- "gsea_kg_nd_frax"
tbl <- gsea_kg_nd$FRAX %>%
    dplyr::select(Pathway, pval, FDR, Edge, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_kg_nd_frax}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_kg_nd$FRAX, FALSE)
```

```{r edge_kg_nd_frax}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_kg_nd$FRAX, FALSE)
```

:::

#### FMR1

::: {.panel-tabset group="nondirectional"}

##### Results table

```{r gsea_kg_nd_fmr1}
id <- "gsea_kg_nd_fmr1"
tbl <- gsea_kg_nd$FMR1 %>%
    dplyr::select(Pathway, pval, FDR, Edge, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_kg_nd_fmr1}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_kg_nd$FMR1, FALSE)
```

```{r edge_kg_nd_fmr1}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_kg_nd$FMR1, FALSE)
```

:::

#### Comparison

```{r heat_kg_nd}
tbl <- gsea_kg_nd$FRAX %>%
    dplyr::select(Pathway, FRAX = NES, sig_frax = Sig) %>%
    left_join(gsea_kg_nd$FMR1) %>%
    dplyr::select(Pathway, FRAX, FMR1 = NES, sig_frax, sig_fmr1 = Sig) %>%
    mutate(
        Pathway = str_trunc(str_replace_all(Pathway, "_", " "), 30),
        sig_any = as.logical(pmax(sig_frax, sig_fmr1))
    ) %>%
    dplyr::filter(sig_any)
mat <- tbl %>%
    dplyr::select(Pathway, FRAX, FMR1) %>%
    column_to_rownames("Pathway") %>%
    as.matrix()
row_anno <-  rowAnnotation(
    Significant = tbl %>%
        mutate(
            category = case_when(
                sig_frax & sig_fmr1 ~ "FRAX & FMR1",
                sig_frax & !sig_fmr1 ~ "FRAX only",
                !sig_frax & sig_fmr1 ~ "FMR1 only"
            ),
            category = fct_relevel(category, c(
                "FRAX only",
                "FMR1 only",
                "FRAX & FMR1"
            ))
        ) %>%
        with(structure(category, names = Pathway)),
    col = list(
        Significant = pal_frax_fmr1
    )
)
Heatmap(
    mat,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 10),
    row_names_side = "left",
    left_annotation = row_anno,
    col = colorRampPalette(c("grey90", "darkblue"))(100),
    heatmap_legend_param = list(
        title = "NES"
    )
)
```

### Wikipathways

```{r gsea_wk_nd}
set.seed(2407)
gsea_wk_nd <- lapply(ranks_nd, \(x){
    fgseaMultilevel(wk_by_id, x, scoreType = "pos") %>%
        as_tibble() %>%
        dplyr::rename(FDR = padj, Pathway = pathway) %>%
        dplyr::arrange(pval) %>%
        mutate(
            Edge = vapply(leadingEdge, length, numeric(1)),
            bonf.padj = p.adjust(pval, "bonferroni"),
            Sig = FDR < 0.05,
        )
})
```

#### FRAX

::: {.panel-tabset group="nondirectional"}

##### Results table

```{r gsea_wk_nd_frax}
id <- "gsea_wk_nd_frax"
tbl <- gsea_wk_nd$FRAX %>%
    dplyr::select(Pathway, pval, FDR, Edge, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_wk_nd_frax}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_wk_nd$FRAX, FALSE)
```

```{r edge_wk_nd_frax}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_wk_nd$FRAX, FALSE)
```

:::

#### FMR1

::: {.panel-tabset group="nondirectional"}

##### Results table

```{r gsea_wk_nd_fmr1}
id <- "gsea_wk_nd_fmr1"
tbl <- gsea_wk_nd$FMR1 %>%
    dplyr::select(Pathway, pval, FDR, Edge, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_wk_nd_fmr1}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_wk_nd$FMR1, FALSE)
```

```{r edge_wk_nd_fmr1}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_wk_nd$FMR1, FALSE)
```

:::

#### Comparison

```{r heat_wk_nd}
tbl <- gsea_wk_nd$FRAX %>%
    dplyr::select(Pathway, FRAX = NES, sig_frax = Sig) %>%
    left_join(gsea_wk_nd$FMR1) %>%
    dplyr::select(Pathway, FRAX, FMR1 = NES, sig_frax, sig_fmr1 = Sig) %>%
    mutate(
        Pathway = str_trunc(str_replace_all(Pathway, "_", " "), 30),
        sig_any = as.logical(pmax(sig_frax, sig_fmr1))
    ) %>%
    dplyr::filter(sig_any)
mat <- tbl %>%
    dplyr::select(Pathway, FRAX, FMR1) %>%
    column_to_rownames("Pathway") %>%
    as.matrix()
row_anno <-  rowAnnotation(
    Significant = tbl %>%
        mutate(
            category = case_when(
                sig_frax & sig_fmr1 ~ "FRAX & FMR1",
                sig_frax & !sig_fmr1 ~ "FRAX only",
                !sig_frax & sig_fmr1 ~ "FMR1 only"
            ),
            category = fct_relevel(category, c(
                "FRAX only",
                "FMR1 only",
                "FRAX & FMR1"
            ))
        ) %>%
        with(structure(category, names = Pathway)),
    col = list(
        Significant = pal_frax_fmr1
    )
)
Heatmap(
    mat,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 10),
    row_names_side = "left",
    left_annotation = row_anno,
    col = colorRampPalette(c("grey90", "darkblue"))(100),
    heatmap_legend_param = list(
        title = "NES"
    )
)
```

### Reactome

```{r gsea_rtm_nd}
set.seed(2407)
gsea_rtm_nd <- lapply(ranks_nd, \(x){
    fgseaMultilevel(rtm_by_id, x, scoreType = "pos") %>%
        as_tibble() %>%
        dplyr::rename(FDR = padj, Pathway = pathway) %>%
        dplyr::arrange(pval) %>%
        mutate(
            Edge = vapply(leadingEdge, length, numeric(1)),
            bonf.padj = p.adjust(pval, "bonferroni"),
            Sig = FDR < 0.05,
        )
})
```

#### FRAX

::: {.panel-tabset group="nondirectional"}

##### Results table

```{r gsea_rtm_nd_frax}
id <- "gsea_rtm_nd_frax"
tbl <- gsea_rtm_nd$FRAX %>%
    dplyr::select(Pathway, pval, FDR, Edge, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_rtm_nd_frax}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_rtm_nd$FRAX, FALSE)
```

```{r edge_rtm_nd_frax}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_rtm_nd$FRAX, FALSE)
```

:::

#### FMR1

::: {.panel-tabset group="nondirectional"}

##### Results table

```{r gsea_rtm_nd_fmr1}
id <- "gsea_rtm_nd_fmr1"
tbl <- gsea_rtm_nd$FMR1 %>%
    dplyr::select(Pathway, pval, FDR, Edge, Sig, leadingEdge) %>%
    mutate(
        Pathway = str_replace_all(Pathway, "_", " "),
        Sig = str_to_title(as.character(Sig)),
        leadingEdge = lapply(leadingEdge, \(x){
            sort(genes$gene_name[genes$gene_id %in% x])
        })
    )

htmltools::browsable(
    tagList(
        tags$caption(glue(
            "{sum(as.logical(tbl$Sig))} of {nrow(tbl)} total pathways ",
            "were classified as significantly enriched. ",
            "This table is interactive where columns can be resized, ",
            "and leading edge genes can be seen by hovering over cells in the 'Edge' column."
        )),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                pval = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e")),
                Pathway = colDef(minWidth = 160),
                Edge = colDef(
                    cell = \(val, ind, col_name){
                        tooltip(
                            val,
                            tbl$leadingEdge[ind]
                        )
                    }
                ),
                leadingEdge = colDef(show = FALSE)
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

##### Network vis

```{r node_rtm_nd_fmr1}
#| fig-cap: "Network analysis of the significantly enriched pathways. Coloured nodes represent pathways, black nodes represent leading edge genes. The relative size of pathway nodes is determined by -log~10~(*p*). Pathways that cluster close together share leading edge genes. Pathway nodes are labelled by their respective name, and in the case of directional GSEA, upregulated pathways are colour blue and downregulated pathways are coloured red."
set.seed(2407)
gsea_node_network(gsea_rtm_nd$FMR1, FALSE)
```

```{r edge_rtm_nd_fmr1}
#| fig-cap: "An alternative representation for network analysis of significantly enriched pathways. Here genes are omitted, but the alpha level (opacity) of edges represents the number of shared genes between the gene sets. Edges that appear bolder relative to other edges indicates that these pathways share a larger number of genes."
set.seed(2407)
gsea_edge_network(gsea_rtm_nd$FMR1, FALSE)
```

:::

#### Comparison

```{r heat_rtm_nd}
#| fig-height: 12
tbl <- gsea_rtm_nd$FRAX %>%
    dplyr::select(Pathway, FRAX = NES, sig_frax = Sig) %>%
    left_join(gsea_rtm_nd$FMR1) %>%
    dplyr::select(Pathway, FRAX, FMR1 = NES, sig_frax, sig_fmr1 = Sig) %>%
    mutate(
        Pathway = str_trunc(str_replace_all(Pathway, "_", " "), 30),
        sig_any = as.logical(pmax(sig_frax, sig_fmr1))
    ) %>%
    dplyr::filter(sig_any)
mat <- tbl %>%
    dplyr::select(Pathway, FRAX, FMR1) %>%
    column_to_rownames("Pathway") %>%
    as.matrix()
row_anno <-  rowAnnotation(
    Significant = tbl %>%
        mutate(
            category = case_when(
                sig_frax & sig_fmr1 ~ "FRAX & FMR1",
                sig_frax & !sig_fmr1 ~ "FRAX only",
                !sig_frax & sig_fmr1 ~ "FMR1 only"
            ),
            category = fct_relevel(category, c(
                "FRAX only",
                "FMR1 only",
                "FRAX & FMR1"
            ))
        ) %>%
        with(structure(category, names = Pathway)),
    col = list(
        Significant = pal_frax_fmr1
    )
)
Heatmap(
    mat,
    show_row_dend = FALSE,
    show_column_dend = FALSE,
    row_names_gp = gpar(fontsize = 7),
    row_names_side = "left",
    left_annotation = row_anno,
    col = colorRampPalette(c("grey90", "darkblue"))(100),
    heatmap_legend_param = list(
        title = "NES"
    )
)
```

# Session information

```{r session_info}
sessionInfo() %>%
    pander()
```
