---
title: "PKTu"
subtitle: "Differential Allelic Representation"
author: "Lachlan Baer"
date-modified: today
date-format: "DD MMMM, YYYY"
modified-title: "LAST MODIFIED"
code-fold: true
fig-width: 8
fig-height: 6
fig-align: center
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    error = FALSE,
    fig.align = "center"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r pkgs}
suppressPackageStartupMessages({
    ## Common packages
    library(tidyverse)
    library(magrittr)
    library(here)
    library(kableExtra)
    library(RColorBrewer)
    library(ggpubr)
    library(scales)
    library(AnnotationHub)
    library(ggrepel)
    library(ggtext)
    library(glue)
    library(pander)
    library(DT)
    library(reactable)
    library(htmltools)
    ## Document-specific packages
    library(tadar)
    library(limma)
    library(plyranges)
})
```

```{r opts}
theme_set(theme_bw())
theme_update(
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown(),
    strip.background = element_blank(),
    strip.text = element_textbox(
        size = 10,
        color = "black", fill = "white", box.color = "black",
        halign = 0.5, linetype = 1, width = unit(1, "npc"),
        padding = margin(3, 0, 1, 0), margin = margin(0, 0, 0, 0)
    )
)
pal1 <- brewer.pal(12, "Set3")
pal2 <- brewer.pal(8, "Set2")
options(ucscChromosomeNames=FALSE)
```

```{r lb_complex_dt}
lb_complex_dt <- function(tbl, cap, rownames = FALSE){
    datatable(
        tbl,
        caption = htmltools::tags$caption(
            cap,
            style = "text-align: left;"
        ),
        options = list(
            dom = "tflipr",  # Available options: tflipr
            columnDefs = list(list(className = 'dt-left', targets = "_all")),  # Align all columns left
            pageLength = 10,  # Show all results
            scrollCollapse = TRUE,  # Removes whitespace when table is small
            scrollY = "100%",  # Sets max height of table before scrollbar is added
            scrollX = TRUE,  # Enable horizontal scrolling when too wide
            sScrollX = "100%",  # Stops last column title being cut off
            initComplete = JS(  # Set font
                "function(settings, json) {",
                "$('body').css({'font-family': 'Calibri'});",
                "}"
            )
        ),
        rownames = rownames,
        height = "100%",  # Seems to work in tandem with scrollCollapse
        width = "100%",  # Fills width when viewed in RStudio viewer
        class = "stripe hover row-border order-column compact nowrap"
    )
}
```

```{r lb_reactable}
lb_reactable <- function(
        tbl, highlight = TRUE, striped = TRUE, compact = TRUE,
        wrap = FALSE, resizable = TRUE, searchable = TRUE,
        style = list(fontFamily = "Calibri, sans-serif"), ...
){
    reactable(
        tbl,
        highlight = highlight, striped = striped, compact = compact,
        wrap = wrap, resizable = TRUE, searchable = TRUE,
        style = style, ...
    )
}
tooltip <- function(value, tooltip) {
    tags$abbr(
        style = "text-decoration: underline; text-decoration-style: solid; cursor: help",
        title = tooltip,
        value
    )
}
react_format <- function(format, digits = 2){
    function(val, ind, col_name){
        formatC(val, digits = digits, format = format)
    }
}
```

# Metadata

```{r meta}
meta <- read_csv(here("misc/metadata.csv")) %>%
    mutate(
        geno_psen1 = fct_relevel(geno_psen1, c("wt", "eofad", "fai")),
        geno_chr14 = fct_relevel(geno_chr14, c("tu", "pk", "het")),
        geno_joint = fct_relevel(geno_joint, unique(geno_joint))
    )
meta %>%
    kable(
        align = "l",
        caption = "Sample metadata"
    ) %>%
    kable_styling(
        bootstrap_options = c("striped", "hover", "condensed", "responsive")
    )
```

```{r pal_group}
pal_group <- c(
    "eofad_tu" = pal1[1], "eofad_pk" = pal1[2], "eofad_het" = pal1[3],
    "fai_tu" = pal1[4], "fai_pk" = pal1[5], "fai_het" = pal1[6],
    "wt_tu" = pal1[7], "wt_pk" = pal1[8], "wt_het" = pal1[10],
    "eofad" = pal2[1], "fai" = pal2[2], "wt" = pal2[3],
    "pk" = pal2[4], "tu" = pal2[5], "het" = pal2[6]
)
```

# Annotations

```{r ensParams}
ens_species <- "Danio rerio"
ens_release <- "111"
ens_assembly <- "GRCz11"
```

```{r ensDb}
ah <- AnnotationHub() %>%
    subset(species == ens_species) %>%
    subset(rdataclass == "EnsDb")
ahId <- ah$ah_id[str_detect(ah$title, ens_release)]
ensDb <- ah[[ahId]]
```

```{r primary_chrs}
primary_chrs <- c(as.character(1:25), "MT")
```

```{r transcripts}
transcripts <- transcripts(ensDb)
tx_len <- exonsBy(ensDb, "tx") %>%
    width() %>%
    lapply(sum) %>%
    unlist(use.names = TRUE) %>%
    enframe(name = "tx_id", value = "length")
mcols(transcripts) <- mcols(transcripts)[
    c("tx_id", "gene_id", "gc_content")
] %>%
    as.data.frame() %>%
    left_join(tx_len)
```

```{r gene_gclen}
gene_gclen <- transcripts %>%
    mcols() %>%
    as_tibble() %>%
    group_by(gene_id) %>%
    summarise(
        gc_content = sum(gc_content*length) / sum(length),
        length = ceiling(median(length))
    )
```

```{r genes}
genes <- genes(ensDb)
mcols(genes) <- mcols(genes)[
    c("gene_id", "gene_name", "gene_biotype", "entrezid")
] %>%
    as.data.frame() %>%
    left_join(gene_gclen)
```

An `EnsDb` object was obtained for Ensembl release `r ens_release` with the `AnnotationHub` package.
This contained the information required to set up gene and transcript annotations.
Gene-level estimates of GC content and length were also generated which are required for assessing potential biases in downstream analysis.

# DE data

```{r}
tt <- readRDS(here("Rdata/tt.Rds"))
```

# Differential Allelic Representation

```{r}
vcf <- here("smk-variants/results/variation/vcf/snvs.final.vcf")
groups <- list(
    eofad = meta$sample[meta$geno_psen1 == "eofad"],
    fai = meta$sample[meta$geno_psen1 == "fai"],
    wt = meta$sample[meta$geno_psen1 == "wt"],
    tu = meta$sample[meta$geno_chr14 == "tu"],
    pk = meta$sample[meta$geno_chr14 == "pk"],
    het = meta$sample[meta$geno_chr14 == "het"]
)
contrasts <- makeContrasts(
    eofad_wt = eofad - wt,
    fai_wt = fai - wt,
    eofad_fai = eofad - fai,
    pk_tu = pk - tu,
    tu_het = tu - het,
    pk_het = pk - het,
    levels = names(groups)
)
```

```{r}
# gts <- tadar::readGenotypes(vcf, genome = genome(genes)) %>%
#     subsetByOverlaps(as(seqinfo(genes), "GRanges")) %>%
#     keepStandardChromosomes(pruning.mode = "coarse")
# props <- gts %>%
#     tadar::countAlleles(groups = groups) %>%
#     tadar::filterLoci(n_called == 18) %>%
#     tadar::countsToProps()
# dar_1e6 <- dar(props, contrasts = contrasts, region_fixed = 1e4)
# saveRDS(dar_1e6, here("Rdata/dar_1e6.Rds"))
# dar_1e6 <- readRDS(here("Rdata/dar_1e6.Rds"))
```

```{r}
# gts <- tadar::readGenotypes(vcf, genome = genome(genes)) %>%
#     subsetByOverlaps(as(seqinfo(genes), "GRanges")) %>%
#     keepStandardChromosomes(pruning.mode = "coarse")
# saveRDS(gts, here("Rdata/tmp/gts.Rds"))
gts <- readRDS(here("Rdata/tmp/gts.Rds"))
```

```{r}
## Calculated DAR group-by-group because R was crashing
# grp <- groups[c("pk", "het")]
# cnt <- makeContrasts(
#     # eofad_wt = eofad - wt,
#     # fai_wt = fai - wt,
#     # eofad_fai = eofad - fai,
#     # pk_tu = pk - tu,
#     # tu_het = tu - het,
#     pk_het = pk - het,
#     levels = names(grp)
# )
# name <- colnames(cnt)
# props <- gts %>%
#     tadar::countAlleles(groups = grp) %>%
#     tadar::filterLoci(n_called == 18) %>%
#     tadar::countsToProps()
# dar_1e6 <- dar(props, contrasts = cnt, region_fixed = 1e6)
# saveRDS(dar_1e6, here(paste0("Rdata/tmp/dar_1e6_", name, ".Rds")))

# dar <- readRDS(here("Rdata/tmp/dar_1e6_eofad_wt.Rds"))
# dar$fai_wt <- readRDS(here("Rdata/tmp/dar_1e6_fai_wt.Rds"))$fai_wt
# dar$eofad_fai <- readRDS(here("Rdata/tmp/dar_1e6_eofad_fai.Rds"))$eofad_fai
# dar$pk_tu <- readRDS(here("Rdata/tmp/dar_1e6_pk_tu.Rds"))$pk_tu
# dar$tu_het <- readRDS(here("Rdata/tmp/dar_1e6_tu_het.Rds"))$tu_het
# dar$pk_het <- readRDS(here("Rdata/tmp/dar_1e6_pk_het.Rds"))$pk_het
# saveRDS(dar, here("Rdata/dar_1e6.Rds"))

dar_1e6 <- readRDS(here("Rdata/dar_1e6.Rds"))
```

```{r}
# gene_dar <- flipRanges(dar) %>%
#     assignFeatureDar(genes, dar_val = "region", fill_missing = 0)
# saveRDS(gene_dar, here("Rdata/gene_dar_1e6.Rds"))
gene_dar <- readRDS(here("Rdata/gene_dar_1e6.Rds"))
```

## ECDF

::: {.panel-tabset}

### EOfAD vs WT

```{r}
plotDarECDF(dar_1e6$eofad_wt, dar_val = "region", highlight = 17)
```

### fAI vs WT

```{r}
plotDarECDF(dar_1e6$fai_wt, dar_val = "region", highlight = 17)
```

### EOfAD vs fAI

```{r}
plotDarECDF(dar_1e6$eofad_fai, dar_val = "region", highlight = 17)
```

### Tu vs PK

```{r}
plotDarECDF(dar_1e6$pk_tu, dar_val = "region", highlight = 14)
```

### Tu vs het

```{r}
plotDarECDF(dar_1e6$tu_het, dar_val = "region", highlight = 14)
```

### PK vs het

```{r}
plotDarECDF(dar_1e6$pk_het, dar_val = "region", highlight = 14)
```

:::

## DEG-DAR

### EOfAD vs WT

```{r}
de_eofad_wt <- tt$eofad_wt %>%
    dplyr::filter(DE) %>%
    pull(gene_id)
```

::: {.panel-tabset}

#### chr1

```{r}
chr <- "1"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr2

```{r}
chr <- "2"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr3

```{r}
chr <- "3"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr4

```{r}
chr <- "4"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr5

```{r}
chr <- "5"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr6

```{r}
chr <- "6"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr7

```{r}
chr <- "7"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr8

```{r}
chr <- "8"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr9

```{r}
chr <- "9"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr10

```{r}
chr <- "10"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr11

```{r}
chr <- "11"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr12

```{r}
chr <- "12"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr13

```{r}
chr <- "13"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr14

```{r}
chr <- "14"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr15

```{r}
chr <- "15"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr16

```{r}
chr <- "16"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr17

```{r}
chr <- "17"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
    )
}
```

#### chr18

```{r}
chr <- "18"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr19

```{r}
chr <- "19"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr20

```{r}
chr <- "20"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr21

```{r}
chr <- "21"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr22

```{r}
chr <- "22"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr23

```{r}
chr <- "23"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr24

```{r}
chr <- "24"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

#### chr25

```{r}
chr <- "25"
de_chr <- plyranges::filter(genes[de_eofad_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_wt, dar_val = "region", chr = chr
    )
}
```

:::

### fAI vs WT

```{r}
de_fai_wt <- tt$fai_wt %>%
    dplyr::filter(DE) %>%
    pull(gene_id)
```

::: {.panel-tabset}

#### chr1

```{r}
chr <- "1"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr2

```{r}
chr <- "2"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr3

```{r}
chr <- "3"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr4

```{r}
chr <- "4"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr5

```{r}
chr <- "5"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr6

```{r}
chr <- "6"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr7

```{r}
chr <- "7"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr8

```{r}
chr <- "8"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr9

```{r}
chr <- "9"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr10

```{r}
chr <- "10"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr11

```{r}
chr <- "11"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr12

```{r}
chr <- "12"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr13

```{r}
chr <- "13"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr14

```{r}
chr <- "14"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr15

```{r}
chr <- "15"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr16

```{r}
chr <- "16"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr17

```{r}
chr <- "17"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
    )
}
```

#### chr18

```{r}
chr <- "18"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr19

```{r}
chr <- "19"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr20

```{r}
chr <- "20"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr21

```{r}
chr <- "21"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr22

```{r}
chr <- "22"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr23

```{r}
chr <- "23"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr24

```{r}
chr <- "24"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

#### chr25

```{r}
chr <- "25"
de_chr <- plyranges::filter(genes[de_fai_wt,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$fai_wt, dar_val = "region", chr = chr
    )
}
```

:::

### EOfAD vs fAI

```{r}
de_eofad_fai <- tt$eofad_fai %>%
    dplyr::filter(DE) %>%
    pull(gene_id)
```

::: {.panel-tabset}

#### chr1

```{r}
chr <- "1"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr2

```{r}
chr <- "2"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr3

```{r}
chr <- "3"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr4

```{r}
chr <- "4"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr5

```{r}
chr <- "5"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr6

```{r}
chr <- "6"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr7

```{r}
chr <- "7"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr8

```{r}
chr <- "8"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr9

```{r}
chr <- "9"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr10

```{r}
chr <- "10"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr11

```{r}
chr <- "11"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr12

```{r}
chr <- "12"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr13

```{r}
chr <- "13"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr14

```{r}
chr <- "14"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr15

```{r}
chr <- "15"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr16

```{r}
chr <- "16"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr17

```{r}
chr <- "17"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
    )
}
```

#### chr18

```{r}
chr <- "18"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr19

```{r}
chr <- "19"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr20

```{r}
chr <- "20"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr21

```{r}
chr <- "21"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr22

```{r}
chr <- "22"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr23

```{r}
chr <- "23"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr24

```{r}
chr <- "24"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

#### chr25

```{r}
chr <- "25"
de_chr <- plyranges::filter(genes[de_eofad_fai,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$eofad_fai, dar_val = "region", chr = chr
    )
}
```

:::

### PK vs Tu

```{r}
de_pk_tu <- tt$pk_tu %>%
    dplyr::filter(DE) %>%
    pull(gene_id)
```

::: {.panel-tabset}

#### chr1

```{r}
chr <- "1"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr2

```{r}
chr <- "2"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr3

```{r}
chr <- "3"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr4

```{r}
chr <- "4"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr5

```{r}
chr <- "5"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr6

```{r}
chr <- "6"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr7

```{r}
chr <- "7"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr8

```{r}
chr <- "8"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr9

```{r}
chr <- "9"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr10

```{r}
chr <- "10"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr11

```{r}
chr <- "11"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr12

```{r}
chr <- "12"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr13

```{r}
chr <- "13"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr14

```{r}
chr <- "14"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr15

```{r}
chr <- "15"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr16

```{r}
chr <- "16"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr17

```{r}
chr <- "17"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
    )
}
```

#### chr18

```{r}
chr <- "18"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr19

```{r}
chr <- "19"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr20

```{r}
chr <- "20"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr21

```{r}
chr <- "21"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr22

```{r}
chr <- "22"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr23

```{r}
chr <- "23"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr24

```{r}
chr <- "24"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

#### chr25

```{r}
chr <- "25"
de_chr <- plyranges::filter(genes[de_pk_tu,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_tu, dar_val = "region", chr = chr
    )
}
```

:::

### Tu vs het

```{r}
de_tu_het <- tt$tu_het %>%
    dplyr::filter(DE) %>%
    pull(gene_id)
```

::: {.panel-tabset}

#### chr1

```{r}
chr <- "1"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr2

```{r}
chr <- "2"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr3

```{r}
chr <- "3"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr4

```{r}
chr <- "4"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr5

```{r}
chr <- "5"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr6

```{r}
chr <- "6"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr7

```{r}
chr <- "7"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr8

```{r}
chr <- "8"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr9

```{r}
chr <- "9"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr10

```{r}
chr <- "10"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr11

```{r}
chr <- "11"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr12

```{r}
chr <- "12"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr13

```{r}
chr <- "13"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr14

```{r}
chr <- "14"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr15

```{r}
chr <- "15"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr16

```{r}
chr <- "16"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr17

```{r}
chr <- "17"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
    )
}
```

#### chr18

```{r}
chr <- "18"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr19

```{r}
chr <- "19"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr20

```{r}
chr <- "20"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr21

```{r}
chr <- "21"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr22

```{r}
chr <- "22"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr23

```{r}
chr <- "23"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr24

```{r}
chr <- "24"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

#### chr25

```{r}
chr <- "25"
de_chr <- plyranges::filter(genes[de_tu_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$tu_het, dar_val = "region", chr = chr
    )
}
```

:::

### PK vs het

```{r}
de_pk_het <- tt$pk_het %>%
    dplyr::filter(DE) %>%
    pull(gene_id)
```

::: {.panel-tabset}

#### chr1

```{r}
chr <- "1"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr2

```{r}
chr <- "2"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr3

```{r}
chr <- "3"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr4

```{r}
chr <- "4"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr5

```{r}
chr <- "5"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr6

```{r}
chr <- "6"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr7

```{r}
chr <- "7"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr8

```{r}
chr <- "8"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr9

```{r}
chr <- "9"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr10

```{r}
chr <- "10"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr11

```{r}
chr <- "11"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr12

```{r}
chr <- "12"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr13

```{r}
chr <- "13"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr14

```{r}
chr <- "14"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr15

```{r}
chr <- "15"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr16

```{r}
chr <- "16"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr17

```{r}
chr <- "17"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        foi = genes[genes$gene_name == "psen1",],
        foi_anno = "gene_name", foi_highlight = TRUE,
    )
}
```

#### chr18

```{r}
chr <- "18"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr19

```{r}
chr <- "19"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr20

```{r}
chr <- "20"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr21

```{r}
chr <- "21"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr22

```{r}
chr <- "22"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr23

```{r}
chr <- "23"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr24

```{r}
chr <- "24"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

#### chr25

```{r}
chr <- "25"
de_chr <- plyranges::filter(genes[de_pk_het,], seqnames == chr)
if (length(de_chr) > 0) {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr,
        features = de_chr,
        features_anno = "gene_name", features_highlight = TRUE
    )
} else {
    plotChrDar(
        dar_1e6$pk_het, dar_val = "region", chr = chr
    )
}
```

:::

# Data export

```{r data_export}

```

# Session information

```{r session_info}
sessionInfo() %>%
    pander()
```