---
title: "PKTu"
subtitle: "Differential Expression Analysis"
author: "Lachlan Baer"
date-modified: today
date-format: "DD MMMM, YYYY"
modified-title: "LAST MODIFIED"
code-fold: true
fig-width: 8
fig-height: 6
fig-align: center
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    error = FALSE,
    fig.align = "center"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r pkgs}
suppressPackageStartupMessages({
    ## Common packages
    library(tidyverse)
    library(magrittr)
    library(here)
    library(kableExtra)
    library(RColorBrewer)
    library(ggpubr)
    library(scales)
    library(AnnotationHub)
    library(ggrepel)
    library(ggtext)
    library(glue)
    library(pander)
    library(DT)
    library(reactable)
    # library(reactablefmtr)
    library(htmltools)
    ## Document-specific packages
    library(edgeR)
    library(cqn)
    library(pheatmap)
    library(corrplot)
})
```

```{r opts}
theme_set(theme_bw())
theme_update(
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown(),
    strip.background = element_blank(),
    strip.text = element_textbox(
        size = 10,
        color = "black", fill = "white", box.color = "black",
        halign = 0.5, linetype = 1, width = unit(1, "npc"),
        padding = margin(3, 0, 1, 0), margin = margin(0, 0, 0, 0)
    )
)
```

```{r lb_reactable}
lb_reactable <- function(
        tbl, highlight = TRUE, striped = TRUE, compact = TRUE,
        wrap = FALSE, resizable = TRUE, searchable = TRUE,
        style = list(fontFamily = "Calibri, sans-serif"), ...
){
    reactable(
        tbl,
        highlight = highlight, striped = striped, compact = compact,
        wrap = wrap, resizable = TRUE, searchable = TRUE,
        style = style, ...
    )
}
tooltip <- function(value, tooltip) {
    tags$abbr(
        style = "text-decoration: underline; text-decoration-style: solid; cursor: help",
        title = tooltip,
        value
    )
}
react_format <- function(format, digits = 2){
    function(val, ind, col_name){
        formatC(val, digits = digits, format = format)
    }
}
```

# Metadata

```{r meta}
meta <- read_csv(here("misc/metadata.csv")) %>%
    mutate(
        geno_psen1 = fct_relevel(geno_psen1, c("wt", "eofad", "fai")),
        geno_chr14 = fct_relevel(geno_chr14, c("tu", "pk", "het")),
        geno_joint = fct_relevel(geno_joint, unique(geno_joint))
    )
meta %>%
    kable(
        align = "l",
        caption = "Sample metadata"
    ) %>%
    kable_styling(
        bootstrap_options = c("striped", "hover", "condensed", "responsive")
    )
```

```{r pal_group}
pal1 <- brewer.pal(12, "Set3")
pal2 <- brewer.pal(8, "Set2")
pal_group <- c(
    "eofad_tu" = pal1[1], "eofad_pk" = pal1[2], "eofad_het" = pal1[3],
    "fai_tu" = pal1[4], "fai_pk" = pal1[5], "fai_het" = pal1[6],
    "wt_tu" = pal1[7], "wt_pk" = pal1[8], "wt_het" = pal1[10],
    "eofad" = pal2[1], "fai" = pal2[2], "wt" = pal2[3],
    "pk" = pal2[4], "tu" = pal2[5], "het" = pal2[6]
) 
```

# Annotations

```{r ensParams}
ens_species <- "Danio rerio"
ens_release <- "111"
ens_assembly <- "GRCz11"
```

```{r ensDb}
ah <- AnnotationHub() %>%
    subset(species == ens_species) %>%
    subset(rdataclass == "EnsDb")
ahId <- ah$ah_id[str_detect(ah$title, ens_release)]
ensDb <- ah[[ahId]]
```

```{r primary_chrs}
primary_chrs <- c(as.character(1:25), "MT")
```

```{r transcripts}
transcripts <- transcripts(ensDb)
tx_len <- exonsBy(ensDb, "tx") %>%
    width() %>%
    lapply(sum) %>%
    unlist(use.names = TRUE) %>%
    enframe(name = "tx_id", value = "length")
mcols(transcripts) <- mcols(transcripts)[
    c("tx_id", "gene_id", "gc_content")
] %>%
    as.data.frame() %>%
    left_join(tx_len)
```

```{r gene_gclen}
gene_gclen <- transcripts %>%
    mcols() %>%
    as_tibble() %>%
    group_by(gene_id) %>%
    summarise(
        gc_content = sum(gc_content*length) / sum(length),
        length = ceiling(median(length))
    )
```

```{r genes}
genes <- genes(ensDb)
mcols(genes) <- mcols(genes)[
    c("gene_id", "gene_name", "gene_biotype", "entrezid")
] %>%
    as.data.frame() %>%
    left_join(gene_gclen)
```

An `EnsDb` object was obtained for Ensembl release `r ens_release` with the `AnnotationHub` package.
This contained the information required to set up gene and transcript annotations.
Gene-level estimates of GC content and length were also generated which are required for assessing potential biases in downstream analysis.

# Counts

```{r counts}
counts <- read_tsv(here("smk-de/results/count/stranded/all.featureCounts"), comment = "#") %>%
    dplyr::select(-c(Chr, Start, End, Strand, Length)) %>%
    set_colnames(basename(colnames(.))) %>%
    set_colnames(str_remove(colnames(.), "\\.bam")) %>%
    as.data.frame() %>%
    column_to_rownames("Geneid")
## Subset for samples of interest
counts <- counts[,meta$sample]
```

```{r dge_list}
min_cpm <- 1
min_samps <- meta %>%
    group_by(geno_joint) %>%
    summarise(n = n()) %>%
    pull(n) %>%
    min()
dge_samples <- tibble(sample = colnames(counts)) %>%
    left_join(meta)
dge_genes <- genes %>%
    as.data.frame() %>%
    .[rownames(counts),] %>%
    dplyr::select(chr = seqnames, everything()) %>%
    mutate(chr = fct_relevel(chr, primary_chrs))
dge_unfilt <- DGEList(
    counts,
    samples = dge_samples,
    genes = dge_genes
) %>%
    calcNormFactors()
keep_genes <- rowSums(cpm(dge_unfilt$counts) >= min_cpm) >= min_samps
dge_list <- dge_unfilt[keep_genes, , keep.lib.sizes=FALSE] %>%
    calcNormFactors()
```

Gene-level count data was loaded from `featureCounts` and imported into `R` as a `DGEList` object.

Genes with low counts across all libraries provide little evidence for differential expression.
The discreteness of these counts also interferes with downstream statistical approximations.
Genes were therefore retained for downstream analysis if they had `r round(min_cpm, digits = 2)` CPM in at least `r min_samps` samples.
<!-- This meant that each remaining gene must have assigned reads detected in at least one genotype group. -->
Filtering is performed independently of which sample belongs to which group so that no bias is introduced.

`r nrow(dge_list)` genes remained for further analysis with library sizes ranging from `r comma(min(dge_list$samples$lib.size))` to `r comma(max(dge_list$samples$lib.size))`.

```{r dens_plots}
#| fig-cap: "*Expression density plots for all samples before and after filtering*"
dens_unfilt <- cpm(dge_unfilt, log = TRUE) %>%
    as.data.frame() %>%
    pivot_longer(
        cols = everything(),
        names_to = "sample",
        values_to = "logCPM"
    ) %>%
    left_join(dge_list$samples) %>%
    ggplot(aes(logCPM, colour = geno_joint, group = sample)) +
    geom_density() +
    scale_colour_manual(values = pal_group) +
    labs(
        title = "Before filtering",
        x = "logCPM",
        y = "Density",
        colour = "Type"
    )
dens_filt <- cpm(dge_list, log = TRUE) %>%
    as.data.frame() %>%
    pivot_longer(
        cols = everything(),
        names_to = "sample",
        values_to = "logCPM"
    ) %>%
    left_join(dge_list$samples) %>%
    ggplot(aes(logCPM, colour = geno_joint, group = sample)) +
    geom_density() +
    scale_colour_manual(values = pal_group) +
    labs(
        title = "After filtering",
        x = "logCPM",
        y = "Density",
        colour = "Type"
    )
ggarrange(dens_unfilt, dens_filt, common.legend = TRUE, legend = "bottom")
```

# PCA

```{r cpm}
cpm <- dge_list %>%
    cpm(log = TRUE)
```

```{r pca}
pca <- cpm %>%
    t() %>%
    prcomp()
pca_vars <- percent_format(0.1)(summary(pca)$importance["Proportion of Variance",])
```

::: {.tabset-panel}

## Joint genotype

```{r pca_plot_joint}
#| fig-cap: "*Principal Component Analysis of gene-level counts*"
pca$x %>%
    as.data.frame() %>%
    rownames_to_column("sample") %>%
    left_join(dge_list$samples) %>%
    as_tibble() %>%
    ggplot(aes(PC1, PC2, colour = geno_joint, fill = geno_joint, label = sample)) +
    geom_point(size = 3) +
    geom_text_repel(show.legend = FALSE) +
    guides(fill = "none") +
    labs(
        title = "Principal component analysis",
        x = paste0("PC1 (", pca_vars[["PC1"]], ")"),
        y = paste0("PC2 (", pca_vars[["PC2"]], ")"),
        colour = "Joint genotype"
    ) +
    scale_colour_manual(values = pal_group)
```

## psen1 genotype

```{r pca_plot_psen1}
#| fig-cap: "*Principal Component Analysis of gene-level counts*"
pca$x %>%
    as.data.frame() %>%
    rownames_to_column("sample") %>%
    left_join(dge_list$samples) %>%
    as_tibble() %>%
    ggplot(aes(PC1, PC2, colour = geno_psen1, fill = geno_psen1, label = sample)) +
    geom_point(size = 3) +
    geom_text_repel(show.legend = FALSE) +
    guides(fill = "none") +
    labs(
        title = "Principal component analysis",
        x = paste0("PC1 (", pca_vars[["PC1"]], ")"),
        y = paste0("PC2 (", pca_vars[["PC2"]], ")"),
        colour = "Type"
    ) +
    scale_colour_manual(values = pal_group)
```

## Chr14 genotype

```{r pca_plot_chr24}
#| fig-cap: "*Principal Component Analysis of gene-level counts*"
pca$x %>%
    as.data.frame() %>%
    rownames_to_column("sample") %>%
    left_join(dge_list$samples) %>%
    as_tibble() %>%
    ggplot(aes(PC1, PC2, colour = geno_chr14, fill = geno_chr14, label = sample)) +
    geom_point(size = 3) +
    geom_text_repel(show.legend = FALSE) +
    guides(fill = "none") +
    labs(
        title = "Principal component analysis",
        x = paste0("PC1 (", pca_vars[["PC1"]], ")"),
        y = paste0("PC2 (", pca_vars[["PC2"]], ")"),
        colour = "Type"
    ) +
    scale_colour_manual(values = pal_group)
```

:::

A Principal Component Analysis (PCA) was performed using the sample logCPM values.
PC1 accounts for a substantial proportion of the variance (`r pca_vars[["PC1"]]`), but PC2 also contains a considerable amount (`r pca_vars[["PC2"]]`).

## PC correlations

```{r gclen_summary}
gclen_summary <- dge_list$counts %>%
    as.data.frame() %>%
    rownames_to_column("gene_id") %>%
    as_tibble() %>%
    pivot_longer(
        cols = colnames(dge_list),
        names_to = "sample",
        values_to = "counts"
    ) %>%
    dplyr::filter(
        counts > 0
    ) %>%
    left_join(
        gene_gclen
    ) %>%
    dplyr::select(
        ends_with("id"), sample, counts, gc_content, length
    ) %>%
    split(f = .$sample) %>%
    lapply(
        function(x){
            DataFrame(
                gc = Rle(x$gc_content, x$counts),
                logLen = Rle(log10(x$length), x$counts)
            )
        }
    ) %>%
    vapply(function(x){
        c(mean(x$gc), sd(x$gc), mean(x$logLen), sd(x$logLen))
    }, numeric(4)
    ) %>%
    t() %>%
    set_colnames(
        c("mn_gc", "sd_gc", "mn_logLen", "sd_logLen")
    ) %>%
    as.data.frame() %>%
    rownames_to_column("sample") %>%
    as_tibble()
```

```{r pca_cor}
pca_cor <- pca$x %>%
    as.data.frame() %>%
    rownames_to_column("sample") %>%
    left_join(gclen_summary) %>%
    as_tibble() %>%
    left_join(dge_list$samples) %>%
    dplyr::select(
        PC1, PC2, PC3,
        mean_gc = mn_gc,
        mean_length = mn_logLen
    ) %>%
    cor()
```

```{r corrplot}
corrplot(
    pca_cor,
    type = "lower",
    diag = FALSE,
    addCoef.col = 1, addCoefasPercent = TRUE,
    tl.col = "black"
)
```

```{r pca_gc_len}
#| fig-height: 16
a <- pca$x %>%
    as.data.frame() %>%
    rownames_to_column("sample") %>%
    left_join(gclen_summary) %>%
    left_join(dge_list$samples) %>%
    as_tibble() %>%
    ggplot(aes(PC1, mn_logLen)) +
    geom_point(aes(colour = geno_joint), size = 3) +
    geom_text_repel(aes(label = sample), show.legend = FALSE) +
    geom_smooth(method = "lm") +
    labs(
        x = paste0("PC1 (", percent(summary(pca)$importance["Proportion of Variance","PC1"]),")"),
        y = "Mean log~10~(Length)",
        colour = "Type"
    ) +
    scale_colour_manual(values = pal_group)
b <- pca$x %>%
    as.data.frame() %>%
    rownames_to_column("sample") %>%
    left_join(gclen_summary) %>%
    left_join(dge_list$samples) %>%
    as_tibble() %>%
    ggplot(aes(PC1, mn_gc)) +
    geom_point(aes(colour = geno_joint), size = 3) +
    geom_text_repel(aes(label = sample), show.legend = FALSE) +
    geom_smooth(method = "lm") +
    # scale_y_continuous(labels = percent) +
    labs(
        x = paste0("PC1 (", percent(summary(pca)$importance["Proportion of Variance","PC1"]),")"),
        y = "Mean GC",
        colour = "Type"
    ) +
    scale_colour_manual(values = pal_group)
ggarrange(
    a, b, nrow = 2,
    common.legend = TRUE, legend = "bottom"
)
```

# Differential expression

## psen1 genotype

### Design

The design matrix was defined containing an intercept representing baseline expression in wildtype samples. 
The remaining two columns represent the presence of either the EofAD or fAI mutant alleles.

```{r design_psen1}
design_psen1 <- model.matrix(~0 + geno_psen1, data = dge_list$samples) %>%
    set_colnames(str_remove(colnames(.), "geno_psen1"))
```

```{r design_plot_psen1}
#| fig-cap: "*Visualisation of the design matrix*"
pheatmap(
    design_psen1[meta$sample,],
    cluster_cols = FALSE,
    cluster_rows = FALSE,
    color = c("white", "grey50"),
    annotation_row = dge_list$samples["geno_psen1"],
    annotation_colors = list(group = pal_group),
    angle_col = "315",
    legend = FALSE
)
```

```{r contrasts_psen1}
contrasts_psen1 <- makeContrasts(
    eofad_wt = eofad - wt,
    fai_wt = fai - wt,
    eofad_fai = eofad - fai,
    levels = colnames(design_psen1)
)
pander(contrasts_psen1)
```

### Model fitting

```{r fit_psen1}
fit_psen1 <- estimateDisp(dge_list, design_psen1) %>%
    glmQLFit()
tt_psen1 <- colnames(contrasts_psen1) %>%
    sapply(function(x){
        glmQLFTest(fit_psen1, contrast = contrasts_psen1[,x]) %>%
            topTags(n = Inf) %>%
            .[["table"]] %>%
            as_tibble() %>%
            arrange(PValue) %>%
            dplyr::select(
                gene_id, gene_name, logFC, logCPM, PValue, FDR, everything()
            ) %>%
            mutate(
                bonfP = p.adjust(PValue, "bonf"),
                DE = ifelse(FDR < 0.05, TRUE, FALSE),
                DE_cat = case_when(
                    logFC <= 0 & DE ~ "Downregulated",
                    logFC >= 0 & DE ~ "Upregulated",
                    !DE ~ "Non-siginificant"
                ),
                coef = x,
                label = glue("<b style='color:{pal_group[coef]}'>{coef}</b>"),
            )
    }, simplify = FALSE)
```

```{r tt_psen1}
tt <- tt_psen1
```

### DEGs

::: {.panel-tabset}

#### EOfAD vs WT

```{r tt_eofad_wt}
id <- "tt_eofad_wt"
tbl <- tt$eofad_wt %>%
    dplyr::filter(DE) %>%
    dplyr::select(gene_id, gene_name, chr, logFC, logCPM, PValue, FDR)

htmltools::browsable(
    tagList(
        tags$caption(glue("The {nrow(tbl)} differentially expressed genes")),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                logFC = colDef(cell = react_format("f")),
                logCPM = colDef(cell = react_format("f")),
                PValue = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e"))
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

```{r de_cat_eofad_wt}
tt$eofad_wt %>%
    pull(DE_cat) %>%
    table() %>%
    pander()
```

#### fAI vs WT

```{r tt_fai_wt}
id <- "tt_fai_wt"
tbl <- tt$fai_wt %>%
    dplyr::filter(DE) %>%
    dplyr::select(gene_id, gene_name, chr, logFC, logCPM, PValue, FDR)

htmltools::browsable(
    tagList(
        tags$caption(glue("The {nrow(tbl)} differentially expressed genes")),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                logFC = colDef(cell = react_format("f")),
                logCPM = colDef(cell = react_format("f")),
                PValue = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e"))
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

```{r de_cat_fai_wt}
tt$fai_wt %>%
    pull(DE_cat) %>%
    table() %>%
    pander()
```

#### EOfAD vs fAI

```{r tt_eofad_fai}
id <- "tt_eofad_fai"
tbl <- tt$eofad_fai %>%
    dplyr::filter(DE) %>%
    dplyr::select(gene_id, gene_name, chr, logFC, logCPM, PValue, FDR)

htmltools::browsable(
    tagList(
        tags$caption(glue("The {nrow(tbl)} differentially expressed genes")),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                logFC = colDef(cell = react_format("f")),
                logCPM = colDef(cell = react_format("f")),
                PValue = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e"))
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

```{r de_cat_eofad_fai}
tt$eofad_fai %>%
    pull(DE_cat) %>%
    table() %>%
    pander()
```

:::

```{r pal_de}
pal_de <- c(`TRUE` = "darkred", `FALSE` = "grey50")
pal_de_cat <- c("Upregulated" = "darkblue", "Downregulated" = "darkred", "Non-significant" = "grey50")
```

### CPM boxplots

Boxplots of the top DE genes and also those that only just reached the significance threshold (FDR < 0.05) to make sure everything looks as expected.

::: {.panel-tabset}

#### EOfAD vs WT

```{r cpm_eofad}
cpm_check <- tt$eofad_wt %>%
    dplyr::filter(DE) %>%
    dplyr::arrange(PValue) %>%
    mutate(label = glue("{gene_id}<br>{gene_name}<br>rank {row_number()}")) %>%
    dplyr::filter(!between(row_number(), 7, n() - 6)) %>%
    dplyr::select(gene_id, label)
cpm(dge_list, log = TRUE) %>%
    .[cpm_check$gene_id, , drop = FALSE] %>%
    as.data.frame() %>%
    rownames_to_column("gene_id") %>%
    pivot_longer(
        !starts_with("gene_id"), names_to = "sample", values_to = "logCPM"
    ) %>%
    left_join(dge_list$samples[c("sample", "geno_psen1")]) %>%
    left_join(dge_list$genes[c("gene_id", "gene_name")]) %>%
    left_join(cpm_check) %>%
    mutate(label = fct_relevel(label, unique(label))) %>%
    dplyr::filter(geno_psen1 %in% c("eofad", "wt")) %>%
    ggplot(aes(geno_psen1, logCPM, fill = geno_psen1)) +
    geom_boxplot() +
    facet_wrap(~label, scales = "free_y", ncol = 3) +
    scale_fill_manual(values = pal_group) +
    labs(x = "Type") +
    theme(legend.position = "none")
```

#### fAI

```{r cpm_fai}
cpm_check <- tt$fai_wt %>%
    dplyr::filter(DE) %>%
    dplyr::arrange(PValue) %>%
    mutate(label = glue("{gene_id}<br>{gene_name}<br>rank {row_number()}")) %>%
    dplyr::filter(!between(row_number(), 7, n() - 6)) %>%
    dplyr::select(gene_id, label)
cpm(dge_list, log = TRUE) %>%
    .[cpm_check$gene_id, , drop = FALSE] %>%
    as.data.frame() %>%
    rownames_to_column("gene_id") %>%
    pivot_longer(
        !starts_with("gene_id"), names_to = "sample", values_to = "logCPM"
    ) %>%
    left_join(dge_list$samples[c("sample", "geno_psen1")]) %>%
    left_join(dge_list$genes[c("gene_id", "gene_name")]) %>%
    left_join(cpm_check) %>%
    mutate(label = fct_relevel(label, unique(label))) %>%
    dplyr::filter(geno_psen1 %in% c("fai", "wt")) %>%
    ggplot(aes(geno_psen1, logCPM, fill = geno_psen1)) +
    geom_boxplot() +
    facet_wrap(~label, scales = "free_y", ncol = 3) +
    scale_fill_manual(values = pal_group) +
    labs(x = "Type") +
    theme(legend.position = "none")
```

#### EOfAD vs fAI

```{r cpm_eofad_fai}
cpm_check <- tt$eofad_fai %>%
    dplyr::filter(DE) %>%
    dplyr::arrange(PValue) %>%
    mutate(label = glue("{gene_id}<br>{gene_name}<br>rank {row_number()}")) %>%
    dplyr::filter(!between(row_number(), 7, n() - 6)) %>%
    dplyr::select(gene_id, label)
cpm(dge_list, log = TRUE) %>%
    .[cpm_check$gene_id, , drop = FALSE] %>%
    as.data.frame() %>%
    rownames_to_column("gene_id") %>%
    pivot_longer(
        !starts_with("gene_id"), names_to = "sample", values_to = "logCPM"
    ) %>%
    left_join(dge_list$samples[c("sample", "geno_psen1")]) %>%
    left_join(dge_list$genes[c("gene_id", "gene_name")]) %>%
    left_join(cpm_check) %>%
    mutate(label = fct_relevel(label, unique(label))) %>%
    dplyr::filter(geno_psen1 %in% c("eofad", "fai")) %>%
    ggplot(aes(geno_psen1, logCPM, fill = geno_psen1)) +
    geom_boxplot() +
    facet_wrap(~label, scales = "free_y", ncol = 3) +
    scale_fill_manual(values = pal_group) +
    labs(x = "Type") +
    theme(legend.position = "none")
```

:::

### Bias checks

#### *p*-value distributions

```{r pdist_psen1}
tt %>%
    bind_rows() %>%
    ggplot(aes(PValue)) +
    geom_histogram(bins = 100, colour = "black", fill = "grey90") +
    facet_wrap(~coef, ncol = 1) +
    labs(
        x = "*p*-value",
        y = "Count"
    )
```

#### Volcano plots

```{r volcano_psen1}
tt %>%
    bind_rows() %>%
    arrange(DE) %>%
    ggplot(aes(logFC, -log10(PValue), colour = DE_cat)) +
    geom_point(alpha = 0.4) +
    geom_text_repel(
        aes(label = gene_name),
        size = 3,
        data = . %>% dplyr::filter(DE,)
    ) +
    facet_wrap(~coef, ncol = 1) +
    scale_colour_manual(values = pal_de_cat) +
    labs(y = "-log~10~(*p*)") +
    theme(legend.position = "none")
```

#### MA plots

```{r ma_psen1}
tt %>%
    bind_rows() %>%
    arrange(DE) %>%
    ggplot(aes(logCPM, logFC)) +
    geom_point(aes(colour = DE_cat), alpha = 0.4) +
    geom_smooth(se = FALSE, colour = "black") +
    facet_wrap(~coef, ncol = 1) +
    scale_colour_manual(values = pal_de_cat) +
    theme(legend.position = "none")
```

#### GC content

```{r gc_bias_psen1}
tt %>%
    bind_rows() %>%
    arrange(DE) %>%
    mutate(stat = -sign(logFC)*log10(PValue)) %>%
    ggplot(aes(gc_content, stat)) +
    geom_point(aes(colour = DE_cat), alpha = 0.4) +
    geom_smooth(se = FALSE) +
    facet_wrap(~coef) +
    geom_text_repel(
        aes(label = gene_name, colour = DE_cat),
        size = 3,
        data = . %>% dplyr::filter(DE)
    ) +
    labs(
        x = "Gene GC content (%)",
        y = "Ranking statistic"
    ) +
    scale_colour_manual(values = pal_de_cat) +
    theme(legend.position = "none")
```

#### Length

```{r len_bias_psen1}
tt %>%
    bind_rows() %>%
    arrange(DE) %>%
    mutate(stat = -sign(logFC)*log10(PValue)) %>%
    ggplot(aes(length, stat)) +
    geom_point(aes(colour = DE_cat), alpha = 0.4) +
    geom_smooth(se = FALSE) +
    facet_wrap(~coef)  +
    geom_text_repel(
        aes(label = gene_name, colour = DE_cat),
        size = 3,
        data = . %>% dplyr::filter(DE)
    ) +
    labs(
        x = "GC length",
        y = "Ranking Statistic"
    ) +
    # coord_cartesian(ylim = c(-10, 10)) +
    scale_x_log10(labels = comma) +
    scale_colour_manual(values = pal_de_cat) +
    theme(legend.position = "none")
```

## Chr14 genotype

### Design

The design matrix was defined containing an intercept representing baseline expression in tu samples.
The remaining two columns represent the presence of either the EofAD or fAI mutant alleles.

```{r design_chr14}
design_chr14 <- model.matrix(~0 + geno_chr14, data = dge_list$samples) %>%
    set_colnames(str_remove(colnames(.), "geno_chr14"))
```

```{r design_plot_chr14}
#| fig-cap: "*Visualisation of the design matrix*"
pheatmap(
    design_chr14[meta$sample,],
    cluster_cols = FALSE,
    cluster_rows = FALSE,
    color = c("white", "grey50"),
    annotation_row = dge_list$samples["geno_chr14"],
    annotation_colors = list(group = pal_group),
    angle_col = "315",
    legend = FALSE
)
```

```{r contrasts_chr14}
contrasts_chr14 <- makeContrasts(
    pk_tu = pk - tu,
    pk_het = pk - het,
    tu_het = tu - het,
    levels = colnames(design_chr14)
)
pander(contrasts_chr14)
```

### Model fitting

```{r fit_chr14}
fit_chr14 <- estimateDisp(dge_list, design_chr14) %>%
    glmQLFit()
tt_chr14 <- colnames(contrasts_chr14) %>%
    sapply(function(x){
        glmQLFTest(fit_chr14, contrast = contrasts_chr14[,x]) %>%
            topTags(n = Inf) %>%
            .[["table"]] %>%
            as_tibble() %>%
            arrange(PValue) %>%
            dplyr::select(
                gene_id, gene_name, logFC, logCPM, PValue, FDR, everything()
            ) %>%
            mutate(
                bonfP = p.adjust(PValue, "bonf"),
                DE = ifelse(FDR < 0.05, TRUE, FALSE),
                DE_cat = case_when(
                    logFC <= 0 & DE ~ "Downregulated",
                    logFC >= 0 & DE ~ "Upregulated",
                    !DE ~ "Non-siginificant"
                ),
                coef = x,
                label = glue("<b style='color:{pal_group[coef]}'>{coef}</b>"),
            )
    }, simplify = FALSE)
```

```{r tt_chr14}
tt$pk_tu <- tt_chr14$pk_tu
tt$pk_het <- tt_chr14$pk_het
tt$tu_het <- tt_chr14$tu_het
```

### DEGs

::: {.panel-tabset}

#### PK vs Tu

```{r tt_pk_tu}
id <- "tt_pk_tu"
tbl <- tt$pk_tu %>%
    dplyr::filter(DE) %>%
    dplyr::select(gene_id, gene_name, chr, logFC, logCPM, PValue, FDR)

htmltools::browsable(
    tagList(
        tags$caption(glue("The {nrow(tbl)} differentially expressed genes")),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                logFC = colDef(cell = react_format("f")),
                logCPM = colDef(cell = react_format("f")),
                PValue = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e"))
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

```{r de_cat_pk_tu}
tt$pk_tu %>%
    pull(DE_cat) %>%
    table() %>%
    pander()
```

#### PK vs het

```{r tt_pk_het}
id <- "tt_pk_het"
tbl <- tt$pk_het %>%
    dplyr::filter(DE) %>%
    dplyr::select(gene_id, gene_name, chr, logFC, logCPM, PValue, FDR)

htmltools::browsable(
    tagList(
        tags$caption(glue("The {nrow(tbl)} differentially expressed genes")),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                logFC = colDef(cell = react_format("f")),
                logCPM = colDef(cell = react_format("f")),
                PValue = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e"))
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

```{r de_cat_pk_het}
tt$pk_het %>%
    pull(DE_cat) %>%
    table() %>%
    pander()
```

#### Tu vs het

```{r tt_tu_het}
id <- "tt_tu_het"
tbl <- tt$tu_het %>%
    dplyr::filter(DE) %>%
    dplyr::select(gene_id, gene_name, chr, logFC, logCPM, PValue, FDR)

htmltools::browsable(
    tagList(
        tags$caption(glue("The {nrow(tbl)} differentially expressed genes")),
        lb_reactable(
            tbl,
            elementId = id,
            defaultColDef = colDef(
                align = "left",
                minWidth = 80
            ),
            columns = list(
                logFC = colDef(cell = react_format("f")),
                logCPM = colDef(cell = react_format("f")),
                PValue = colDef(cell = react_format("e")),
                FDR = colDef(cell = react_format("e"))
            )
        ),
        tags$button(
            tagList(fontawesome::fa("download"), "Download as CSV"),
            onclick = glue("Reactable.downloadDataCSV('{id}', '{id}.csv')")
        )
    )
)
```

```{r de_cat_tu_het}
tt$tu_het %>%
    pull(DE_cat) %>%
    table() %>%
    pander()
```

:::

### CPM boxplots

Boxplots of the top DE genes and also those that only just reached the significance threshold (FDR < 0.05) to make sure everything looks as expected.

::: {.panel-tabset}

#### PK vs Tu

```{r cpm_pk_tu}
cpm_check <- tt$pk_tu %>%
    dplyr::filter(DE) %>%
    dplyr::arrange(PValue) %>%
    mutate(label = glue("{gene_id}<br>{gene_name}<br>rank {row_number()}")) %>%
    dplyr::filter(!between(row_number(), 7, n() - 6)) %>%
    dplyr::select(gene_id, label)
cpm(dge_list, log = TRUE) %>%
    .[cpm_check$gene_id, , drop = FALSE] %>%
    as.data.frame() %>%
    rownames_to_column("gene_id") %>%
    pivot_longer(
        !starts_with("gene_id"), names_to = "sample", values_to = "logCPM"
    ) %>%
    left_join(dge_list$samples[c("sample", "geno_chr14")]) %>%
    left_join(dge_list$genes[c("gene_id", "gene_name")]) %>%
    left_join(cpm_check) %>%
    mutate(label = fct_relevel(label, unique(label))) %>%
    dplyr::filter(geno_chr14 %in% c("pk", "tu")) %>%
    ggplot(aes(geno_chr14, logCPM, fill = geno_chr14)) +
    geom_boxplot() +
    facet_wrap(~label, scales = "free_y", ncol = 3) +
    scale_fill_manual(values = pal_group) +
    labs(x = "Type") +
    theme(legend.position = "none")
```

#### PK vs het

```{r cpm_pk_het}
cpm_check <- tt$pk_het %>%
    dplyr::filter(DE) %>%
    dplyr::arrange(PValue) %>%
    mutate(label = glue("{gene_id}<br>{gene_name}<br>rank {row_number()}")) %>%
    dplyr::filter(!between(row_number(), 7, n() - 6)) %>%
    dplyr::select(gene_id, label)
cpm(dge_list, log = TRUE) %>%
    .[cpm_check$gene_id, , drop = FALSE] %>%
    as.data.frame() %>%
    rownames_to_column("gene_id") %>%
    pivot_longer(
        !starts_with("gene_id"), names_to = "sample", values_to = "logCPM"
    ) %>%
    left_join(dge_list$samples[c("sample", "geno_chr14")]) %>%
    left_join(dge_list$genes[c("gene_id", "gene_name")]) %>%
    left_join(cpm_check) %>%
    mutate(label = fct_relevel(label, unique(label))) %>%
    dplyr::filter(geno_chr14 %in% c("pk", "het")) %>%
    ggplot(aes(geno_chr14, logCPM, fill = geno_chr14)) +
    geom_boxplot() +
    facet_wrap(~label, scales = "free_y", ncol = 3) +
    scale_fill_manual(values = pal_group) +
    labs(x = "Type") +
    theme(legend.position = "none")
```

#### Tu vs het

```{r cpm_tu_het}
# cpm_check <- tt$tu_het %>%
#     dplyr::filter(DE) %>%
#     dplyr::arrange(PValue) %>%
#     mutate(label = glue("{gene_id}<br>{gene_name}<br>rank {row_number()}")) %>%
#     dplyr::filter(!between(row_number(), 7, n() - 6)) %>%
#     dplyr::select(gene_id, label)
# cpm(dge_list, log = TRUE) %>%
#     .[cpm_check$gene_id, , drop = FALSE] %>%
#     as.data.frame() %>%
#     rownames_to_column("gene_id") %>%
#     pivot_longer(
#         !starts_with("gene_id"), names_to = "sample", values_to = "logCPM"
#     ) %>%
#     left_join(dge_list$samples[c("sample", "geno_chr14")]) %>%
#     left_join(dge_list$genes[c("gene_id", "gene_name")]) %>%
#     left_join(cpm_check) %>%
#     mutate(label = fct_relevel(label, unique(label))) %>%
#     dplyr::filter(geno_chr14 %in% c("tu", "het")) %>%
#     ggplot(aes(geno_chr14, logCPM, fill = geno_chr14)) +
#     geom_boxplot() +
#     facet_wrap(~label, scales = "free_y", ncol = 3) +
#     scale_fill_manual(values = pal_group) +
#     labs(x = "Type") +
#     theme(legend.position = "none")
```

:::

### Bias checks

#### *p*-value distributions

```{r pdist_chr14}
tt %>%
    bind_rows() %>%
    ggplot(aes(PValue)) +
    geom_histogram(bins = 100, colour = "black", fill = "grey90") +
    facet_wrap(~coef, ncol = 1) +
    labs(
        x = "*p*-value",
        y = "Count"
    )
```

#### Volcano plots

```{r volcano_chr14}
#| fig-height: 12
tt %>%
    bind_rows() %>%
    arrange(DE) %>%
    ggplot(aes(logFC, -log10(PValue), colour = DE_cat)) +
    geom_point(alpha = 0.4) +
    geom_text_repel(
        aes(label = gene_name),
        size = 3,
        data = . %>% dplyr::filter(DE,)
    ) +
    facet_wrap(~coef, ncol = 1) +
    scale_colour_manual(values = pal_de_cat) +
    labs(y = "-log~10~(*p*)") +
    theme(legend.position = "none")
```

#### MA plots

```{r ma_chr14}
#| fig-height: 12
tt %>%
    bind_rows() %>%
    arrange(DE) %>%
    ggplot(aes(logCPM, logFC)) +
    geom_point(aes(colour = DE_cat), alpha = 0.4) +
    geom_smooth(se = FALSE, colour = "black") +
    facet_wrap(~coef, ncol = 1) +
    scale_colour_manual(values = pal_de_cat) +
    theme(legend.position = "none")
```

#### GC content

```{r gc_bias_chr14}
tt %>%
    bind_rows() %>%
    arrange(DE) %>%
    mutate(stat = -sign(logFC)*log10(PValue)) %>%
    ggplot(aes(gc_content, stat)) +
    geom_point(aes(colour = DE_cat), alpha = 0.4) +
    geom_smooth(se = FALSE) +
    facet_wrap(~coef) +
    geom_text_repel(
        aes(label = gene_name, colour = DE_cat),
        size = 3,
        data = . %>% dplyr::filter(DE)
    ) +
    labs(
        x = "Gene GC content (%)",
        y = "Ranking statistic"
    ) +
    scale_colour_manual(values = pal_de_cat) +
    theme(legend.position = "none")
```

#### Length

```{r len_bias_chr14}
tt %>%
    bind_rows() %>%
    arrange(DE) %>%
    mutate(stat = -sign(logFC)*log10(PValue)) %>%
    ggplot(aes(length, stat)) +
    geom_point(aes(colour = DE_cat), alpha = 0.4) +
    geom_smooth(se = FALSE) +
    facet_wrap(~coef)  +
    geom_text_repel(
        aes(label = gene_name, colour = DE_cat),
        size = 3,
        data = . %>% dplyr::filter(DE)
    ) +
    labs(
        x = "GC length",
        y = "Ranking Statistic"
    ) +
    # coord_cartesian(ylim = c(-10, 10)) +
    scale_x_log10(labels = comma) +
    scale_colour_manual(values = pal_de_cat) +
    theme(legend.position = "none")
```

# Data export

```{r data_export}
saveRDS(dge_list, here("Rdata/dge_list.Rds"))
saveRDS(tt, here("Rdata/tt.Rds"))
```

# Session information

```{r session_info}
sessionInfo() %>%
    pander()
```