---
title: "PKTu"
subtitle: "Differential Expression Analysis (CQN)"
author: "Lachlan Baer"
date-modified: today
date-format: "DD MMMM, YYYY"
modified-title: "LAST MODIFIED"
code-fold: true
fig-width: 8
fig-height: 6
fig-align: center
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    error = FALSE,
    fig.align = "center"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r pkgs}
suppressPackageStartupMessages({
    ## Common packages
    library(tidyverse)
    library(magrittr)
    library(here)
    library(kableExtra)
    library(RColorBrewer)
    library(ggpubr)
    library(scales)
    library(AnnotationHub)
    library(ggrepel)
    library(ggtext)
    library(glue)
    library(pander)
    library(DT)
    library(reactable)
    # library(reactablefmtr)
    library(htmltools)
    ## Document-specific packages
    library(edgeR)
    library(cqn)
    library(pheatmap)
    library(corrplot)
})
```

```{r opts}
theme_set(theme_bw())
theme_update(
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown(),
    strip.background = element_blank(),
    strip.text = element_textbox(
        size = 10,
        color = "black", fill = "white", box.color = "black",
        halign = 0.5, linetype = 1, width = unit(1, "npc"),
        padding = margin(3, 0, 1, 0), margin = margin(0, 0, 0, 0)
    )
)
pal1 <- brewer.pal(12, "Set3")
pal2 <- brewer.pal(8, "Set2")
```

```{r lb_complex_dt}
lb_complex_dt <- function(tbl, cap, rownames = FALSE){
    datatable(
        tbl,
        caption = htmltools::tags$caption(
            cap,
            style = "text-align: left;"
        ),
        options = list(
            dom = "tflipr",  # Available options: tflipr
            columnDefs = list(list(className = 'dt-left', targets = "_all")),  # Align all columns left
            pageLength = 10,  # Show all results
            scrollCollapse = TRUE,  # Removes whitespace when table is small
            scrollY = "100%",  # Sets max height of table before scrollbar is added
            scrollX = TRUE,  # Enable horizontal scrolling when too wide
            sScrollX = "100%",  # Stops last column title being cut off
            initComplete = JS(  # Set font
                "function(settings, json) {",
                "$('body').css({'font-family': 'Calibri'});",
                "}"
            )
        ),
        rownames = rownames,
        height = "100%",  # Seems to work in tandem with scrollCollapse
        width = "100%",  # Fills width when viewed in RStudio viewer
        class = "stripe hover row-border order-column compact nowrap"
    )
}
```

```{r lb_reactable}
lb_reactable <- function(
        tbl, highlight = TRUE, striped = TRUE, compact = TRUE,
        wrap = FALSE, resizable = TRUE, searchable = TRUE,
        style = list(fontFamily = "Calibri, sans-serif"), ...
){
    reactable(
        tbl,
        highlight = highlight, striped = striped, compact = compact,
        wrap = wrap, resizable = TRUE, searchable = TRUE,
        style = style, ...
    )
}
tooltip <- function(value, tooltip) {
    tags$abbr(
        style = "text-decoration: underline; text-decoration-style: solid; cursor: help",
        title = tooltip,
        value
    )
}
react_format <- function(format, digits = 2){
    function(val, ind, col_name){
        formatC(val, digits = digits, format = format)
    }
}
```

# Metadata

```{r meta}
meta <- read_csv(here("misc/metadata.csv")) %>%
    mutate(
        geno_psen1 = fct_relevel(geno_psen1, c("wt", "eofad", "fai")),
        geno_chr14 = fct_relevel(geno_chr14, c("tu", "pk", "het")),
        geno_joint = fct_relevel(geno_joint, unique(geno_joint))
    )
meta %>%
    kable(
        align = "l",
        caption = "Sample metadata"
    ) %>%
    kable_styling(
        bootstrap_options = c("striped", "hover", "condensed", "responsive")
    )
```

```{r pal_group}
pal_group <- c(
    "eofad_tu" = pal1[1], "eofad_pk" = pal1[2], "eofad_het" = pal1[3],
    "fai_tu" = pal1[4], "fai_pk" = pal1[5], "fai_het" = pal1[6],
    "wt_tu" = pal1[7], "wt_pk" = pal1[8], "wt_het" = pal1[10],
    "eofad" = pal2[1], "fai" = pal2[2], "wt" = pal2[3],
    "pk" = pal2[4], "tu" = pal2[5], "het" = pal2[6]
) 
```

# Annotations

```{r ensParams}
ens_species <- "Danio rerio"
ens_release <- "111"
ens_assembly <- "GRCz11"
```

```{r ensDb}
ah <- AnnotationHub() %>%
    subset(species == ens_species) %>%
    subset(rdataclass == "EnsDb")
ahId <- ah$ah_id[str_detect(ah$title, ens_release)]
ensDb <- ah[[ahId]]
```

```{r primary_chrs}
primary_chrs <- c(as.character(1:25), "MT")
```

```{r transcripts}
transcripts <- transcripts(ensDb)
tx_len <- exonsBy(ensDb, "tx") %>%
    width() %>%
    lapply(sum) %>%
    unlist(use.names = TRUE) %>%
    enframe(name = "tx_id", value = "length")
mcols(transcripts) <- mcols(transcripts)[
    c("tx_id", "gene_id", "gc_content")
] %>%
    as.data.frame() %>%
    left_join(tx_len)
```

```{r gene_gclen}
gene_gclen <- transcripts %>%
    mcols() %>%
    as_tibble() %>%
    group_by(gene_id) %>%
    summarise(
        gc_content = sum(gc_content*length) / sum(length),
        length = ceiling(median(length))
    )
```

```{r genes}
genes <- genes(ensDb)
mcols(genes) <- mcols(genes)[
    c("gene_id", "gene_name", "gene_biotype", "entrezid")
] %>%
    as.data.frame() %>%
    left_join(gene_gclen)
```

An `EnsDb` object was obtained for Ensembl release `r ens_release` with the `AnnotationHub` package.
This contained the information required to set up gene and transcript annotations.
Gene-level estimates of GC content and length were also generated which are required for assessing potential biases in downstream analysis.

# Load data

```{r load_data}
dge_list <- readRDS(here("Rdata/dge_list.Rds"))
tt <- readRDS(here("Rdata/tt.Rds"))
```

# Genotype checks

```{r}
cpm(dge_list, log = TRUE) %>%
    .["ENSDARG00000004870",] %>%
    t() %>%
    data.frame() %>%
    set_rownames("ENSDARG00000004870") %>%
    as.data.frame() %>%
    rownames_to_column("gene_id") %>%
    pivot_longer(
        !starts_with("gene_id"), names_to = "sample", values_to = "logCPM"
    ) %>%
    left_join(dge_list$samples[c("sample", "geno_psen1")]) %>%
    left_join(dge_list$genes[c("gene_id", "gene_name")]) %>%
    mutate(gene_name = factor(gene_name, levels = unique(gene_name))) %>%
    ggplot(aes(sample, logCPM, fill = geno_psen1)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = pal_group) +
    labs(x = "Sample", fill = "*psen1* genotype") +
    theme(legend.title = element_markdown())
```

# Session information

```{r session_info}
sessionInfo() %>%
    pander()
```