---
title: "PKTu"
subtitle: "Quality control"
author: "Lachlan Baer"
date-modified: today
date-format: "DD MMMM, YYYY"
modified-title: "LAST MODIFIED"
code-fold: true
fig-width: 8
fig-height: 8
fig-align: center
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    error = FALSE,
    fig.align = "center"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r pkgs}
suppressPackageStartupMessages({
    ## Common packages
    library(tidyverse)
    library(magrittr)
    library(here)
    library(kableExtra)
    library(RColorBrewer)
    library(ggpubr)
    library(scales)
    library(AnnotationHub)
    library(ggrepel)
    library(ggtext)
    library(glue)
    library(DT)
    library(pander)
    ## Document-specific packages
    library(ngsReports)
})
```

```{r opts}
theme_set(theme_bw())
```

# Quality control

This document outlines the quality control checks that were performed on data at various stages of the pre-processing workflow.

The procedure for generating gene-level counts from raw RNA-seq data can be summarised as: 

<center>
raw reads (.fastq.gz) $\rightarrow$ trimmed reads (.fastq.gz) $\rightarrow$ aligned reads (.bam) $\rightarrow$ counts (.tsv)
</center>

# Metadata

```{r meta}
meta <- read_csv(here("misc/metadata.csv"))
meta %>%
    kable(
        align = "l",
        caption = "Sample metadata"
    ) %>%
    kable_styling(
        bootstrap_options = c("striped", "hover", "condensed", "responsive")
    )
```

# FastQC reports

## Load data

```{r fqc_raw}
#| results: hide
fqc_raw <- list.files(
    path = here("smk-de/results/raw_data/FastQC"),
    pattern = "zip",
    full.names = TRUE
)
## Fix raw FQC zip file structure caused by workflow renaming of FQC files
## Otherwise we encounter a bug in ngsReports
fqc_raw %>%
    lapply(\(x){
        contents <- unzip(x, list = TRUE)
        old_dir <- str_remove(contents$Name[1], "/$")
        new_dir <- str_remove(basename(x), ".zip$")
        if (old_dir != new_dir) {
            unzip(x)
            file.remove(x)
            system(paste('mv', old_dir, new_dir))
            files_to_zip <- list.files(here(old_dir), full.names = TRUE)
            zip(x, new_dir)
            unlink(new_dir, recursive = TRUE)
        }
    })
fqc_raw <- fqc_raw %>%
    FastqcDataList()
## Rename fqc_raw to names used by workflow
sample_rename <- read_tsv(here("smk-de/config/units.tsv")) %>%
    pivot_longer(cols = c("fq1", "fq2"), values_to = "fq") %>%
    dplyr::select(new = sample, old = fq) %>%
    mutate(
        old = basename(old),
        new = ifelse(
            str_detect(old, "_R1_001.fastq.gz$"),
            paste0(new, "_1_R1.fastq.gz"),
            paste0(new, "_1_R2.fastq.gz")
        )
    )
fqName(fqc_raw) <- vapply(fqName(fqc_raw), \(x){
    sample_rename$new[sample_rename$old == x]
}, character(1), USE.NAMES = FALSE)
## Subset for samples we want to analyse
keep_raw <- vapply(fqName(fqc_raw), \(x){
    str_remove(x, "_1_R(1|2).fastq.gz") %in% meta$sample
}, logical(1))
fqc_raw <- fqc_raw[keep_raw]
```

```{r fqc_trim}
fqc_trim <- list.files(
    path = here("smk-de/results/trim/FastQC"),
    pattern = "zip",
    full.names = TRUE
) %>%
    FastqcDataList()
## Subset for samples we want to analyse
keep_trim <- vapply(fqName(fqc_trim), \(x){
    str_remove(x, "_1_R(1|2).fastq.gz") %in% meta$sample
}, logical(1))
fqc_trim <- fqc_trim[keep_trim]
```

```{r fqc_align}
fqc_align <- list.files(
    path = here("smk-de/results/align/FastQC"),
    pattern = "zip",
    full.names = TRUE
) %>%
    FastqcDataList()
## Subset for samples we want to analyse
keep_align <- vapply(fqName(fqc_align), \(x){
    str_remove(x, ".bam") %in% meta$sample
}, logical(1))
fqc_align <- fqc_align[keep_align]
```

The following plots were produced directly from FastQC zip archives with the `ngsReports` package.

Direct comparisons between different stages of pre-processing can be made by toggling the tabs above the figures.

## Read totals

::: {.panel-tabset}

### Raw

```{r read_totals_raw}
plotReadTotals(fqc_raw)
```

### Trimmed

```{r read_totals_trim}
plotReadTotals(fqc_trim)
```

### Aligned

```{r read_totals_align}
plotReadTotals(fqc_align)
```

:::

## FastQC summary

::: {.panel-tabset}

### Raw

```{r summary_raw}
plotSummary(fqc_raw)
```

### Trimmed

```{r summary_trim}
plotSummary(fqc_trim)
```

### Aligned

```{r summary_align}
plotSummary(fqc_align)
```

:::

FastQC reports are well-known to return WARN or FAIL when there is no actual reason for concern.

We assess each QC category individually in the following plots.

## Sequence length distribution

::: {.panel-tabset}

### Raw

```{r seq_len_distn_raw}
plotSeqLengthDistn(fqc_raw)
```

### Trimmed

```{r seq_len_distn_trim}
plotSeqLengthDistn(fqc_trim)
```

### Aligned

```{r seq_len_distn_align}
plotSeqLengthDistn(fqc_align)
```

:::


## Per base sequence qualities

::: {.panel-tabset}

### Raw

```{r base_quals_raw}
plotBaseQuals(fqc_raw)
```

### Trimmed

```{r base_quals_trim}
plotBaseQuals(fqc_trim)
```

### Aligned

```{r base_quals_align}
plotBaseQuals(fqc_align)
```

:::

## Mean sequence quality per read

::: {.panel-tabset}

### Raw

```{r seq_quals_raw}
plotSeqQuals(fqc_raw)
```

### Trimmed

```{r seq_quals_trim}
plotSeqQuals(fqc_trim)
```

### Aligned

```{r seq_quals_align}
plotSeqQuals(fqc_align)
```

:::

## Per base sequence content

::: {.panel-tabset}

### Raw

```{r seq_content_raw}
# plotSeqContent(fqc_raw)
```

### Trimmed

```{r seq_content_trim}
# plotSeqContent(fqc_trim)
```

### Aligned

```{r seq_content_align}
# plotSeqContent(fqc_align)
```

:::

This plotting function appears to be bugged.

## Adapter content

::: {.panel-tabset}

### Raw

```{r adapter_content_raw}
plotAdapterContent(fqc_raw)
```

### Trimmed

```{r adapter_content_trim}
plotAdapterContent(fqc_trim)
```

### Aligned

```{r adapter_content_align}
plotAdapterContent(fqc_align)
```

:::

Trimming successfully remove adapter content.

## Sequence duplication levels

::: {.panel-tabset}

### Raw

```{r dup_levels_raw}
plotDupLevels(fqc_raw)
```

### Trimmed

```{r dup_levels_trim}
plotDupLevels(fqc_trim)
```

### Aligned

```{r dup_levels_align}
plotDupLevels(fqc_align)
```

:::

## GC content

::: {.panel-tabset}

### Raw

```{r gc_content_raw}
plotGcContent(fqc_raw)
```

### Trimmed

```{r gc_content_trim}
plotGcContent(fqc_trim)
```

### Aligned

```{r gc_content_align}
plotGcContent(fqc_align)
```

:::

::: {.panel-tabset}

### Raw

```{r gc_content_2_raw}
#| fig-height: 6
plotGcContent(
    fqc_raw, plotType = "line",  gcType = "Transcriptome", usePlotly = TRUE
)
```

### Trimmed

```{r gc_content_2_trim}
#| fig-height: 6
plotGcContent(
    fqc_trim, plotType = "line",  gcType = "Transcriptome", usePlotly = TRUE
)
```

### Aligned

```{r gc_content_2_align}
#| fig-height: 6
plotGcContent(
    fqc_align, plotType = "line",  gcType = "Transcriptome", usePlotly = TRUE
)
```

:::

Note that the FRAX_294 sample shows a slightly different GC profile to all other samples.

# Alignment statistics

```{r star_logs}
star_logs <- list.files(
    here("smk-de/results/align/log"),
    pattern = ".log.final.out",
    full.names = TRUE
) %>%
    importNgsLogs(type = "star") %>%
    mutate(Filename = str_remove(Filename, ".log.final.out"))
```

```{r alignment_rate}
star_logs %>%
    dplyr::select(Filename, Total_Mapped_Percent) %>%
    mutate(Total_Mapped_Percent = Total_Mapped_Percent / 100) %>%
    ggplot(aes(Filename, Total_Mapped_Percent)) +
    geom_bar(stat = "identity", fill = brewer.pal(8, "Set2")[1]) +
    scale_y_continuous(
        labels = scales::percent,
        breaks = seq(0, 1, 0.1),
        limits = c(0, 1),
        expand = expansion(c(0, 0.02))
    ) +
    coord_flip() +
    labs(
        title = "Read alignment rate",
        x = "Sample",
        y = "Percent mapped reads",
    )
```

```{r alignment_summary}
star_logs %>%
    dplyr::select(
        Filename,
        Mapped_To_Multiple_Loci = Number_Of_Reads_Mapped_To_Multiple_Loci,
        Mapped_To_Too_Many_Loci = Number_Of_Reads_Mapped_To_Too_Many_Loci,
        Unmapped_Other = Number_Of_Reads_Unmapped_Other,
        Uniquely_Mapped = Uniquely_Mapped_Reads_Number
    ) %>%
    pivot_longer(
        cols = c(
            "Mapped_To_Multiple_Loci", "Mapped_To_Too_Many_Loci",
            "Unmapped_Other", "Uniquely_Mapped"
        ),
        names_to = "Type",
        values_to = "number"
    ) %>%
    mutate(
        Type = str_to_sentence(str_replace_all(Type, "_", " ")),
        Type = factor(Type, levels = c(
            "Unmapped other", "Mapped to too many loci",
            "Mapped to multiple loci", "Uniquely mapped"
        ))
    ) %>%
    ggplot(aes(x = Filename, y = number, fill = Type)) +
    geom_bar(position = "stack", stat = "identity") +
    coord_flip() +
    scale_fill_manual(values = brewer.pal(4, "Set2")[4:1]) +
    labs(
        title = "Alignment summary",
        x = "Sample",
        y = "Total reads"
    ) +
    scale_y_continuous(
        labels = scales::comma,
        breaks = seq(0, 1e9, 2.5e7),
        expand = expansion(c(0, 0.09))
    ) +
    theme(legend.position = "bottom")
```

# Gene count statistics

Gene-level counts were determined with `featureCounts` from the `subread` package.

Gene count data was transferred to this project directory manually from the HPC system.
The count output is ~50MB, which is bordering the size limit for file uploads to GitHub.
The snakemake workflow in this repository is however available should these files need to be reproduced.

```{r load_fc_summary}
load_fc_summary <- function(file){
    read_tsv(file) %>%
        pivot_longer(cols = ends_with(".bam"), names_to = "sample", values_to = "reads") %>%
        mutate(
            Status = str_replace_all(Status, "_", " "),
            Status = factor(Status),
            sample = str_remove(basename(sample), ".bam$"),
            sample = factor(sample),
        )
}
```

```{r fc}
fc <- load_fc_summary(here("smk-de/results/count/all.featureCounts.summary"))
```

Description of `featureCounts` unassigned read types:

- Unassigned Unmapped: unmapped reads cannot be assigned.
- Unassigned Read Type: reads that have an unexpected read type (eg. being a single end read included in a paired end dataset) and also cannot be counted with confidence (eg. due to stranded counting). Such reads are typically generated from a read trimming program.
- Unassigned Singleton: read pairs that have only one end mapped.
- Unassigned MappingQuality: alignments with a mapping quality score lower than the threshold.
- Unassigned Chimera: two ends in a paired end alignment are located on different chromosomes or have unexpected orientation.
- Unassigned FragementLength: fragment length inferred from paired end alignment does not meet the length criteria.
- Unassigned Duplicate: alignments marked as duplicate (indicated in the FLAG field).
- Unassigned MultiMapping: alignments reported for multi-mapping reads (indicated by ‘NH’ tag).
- Unassigned Secondary: alignments reported as secondary alignments (indicated in the FLAG field).
- Unassigned Split (or Unassigned NonSplit): alignments that contain junctions (or do not contain junctions).
- Unassigned NoFeatures: alignments that do not overlap any feature.
- Unassigned Overlapping Length: alignments that do not overlap any feature (or metafeature) with the minimum required overlap length.
- Unassigned Ambiguity: alignments that overlap two or more features (feature-level summarization) or meta-features (meta-feature-level summarization).

```{r counts_summary}
fc %>%
    dplyr::filter(reads != 0) %>%
    ggplot(aes(x = sample, y = reads, fill = Status)) +
    geom_bar(position = "stack", stat = "identity") +
    coord_flip() +
    scale_fill_manual(values = brewer.pal(8, "Set2")) +
    labs(
        title = "Gene counts summary",
        x = "Sample",
        y = "Total reads",
        fill = "Type"
    ) +
    scale_y_continuous(
        labels = scales::comma,
        breaks = seq(0, 5e8, 2.5e7),
        expand = expansion(c(0, 0.01))
    ) +
    theme(
        axis.text.x = element_text(angle = -40, , vjust = 1, hjust = 0)
    )
```
